---
title: "Krill TSPIN and EPAC Bongo Biomass distribution models - OPAL"
author: "Solene Derville"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
   html_document:
    highlight: pygments
    keep_md: yes
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
---

### FILE PATHS
```{r, echo=F, result="hide", include=F}
lib=lapply(c("plyr","officer","raster","stars", "sf", "viridis", "ggthemes", "patchwork", "doParallel", "foreach", 
             "tidyverse", "lubridate", "mgcv", "gratia", "ape", "gstat", "RColorBrewer", "caret", "gbm", "dismo", 
             "pROC", "gstat", "future", "future.apply", "furrr", "fmsb", "gridExtra"), library, character.only=T)

mon_theme <- theme(
        panel.border = element_rect(linewidth=0.5,color="black", fill="transparent"),
        plot.margin=unit(c(2,2,2,2),"mm"),
        panel.background = element_rect(fill = 'white'),
        text=element_text(face="bold", size=7),
        title = element_text(size=rel(1)))

krillsp_col <- data.frame(group = c("EPAC", "TSPIN"),
                          col = c("#e1c340", "#4cd7d0"))
krillsp_pal <- c("#e1c340",  "#4cd7d0")
```


Environmental data for predictions
```{r, warning = F}
load("../../2-Environment/Outputs/env_pred_full.RData")
depth_ras <- env_pred_full$env_ras[[1]]$data_pred[[1]]$DEPTH
```

Environmental data for maps
```{r, warning = F, message = F}
# coastline shapefile
coast_sf <- st_read("../../Data/Environment/coast_openstreetmap_WA-OR-CA.shp")
coast_sf_utm <- st_transform(coast_sf, crs = 32610)

# isobaths
iso_sf <- st_read("../../Data/Environment/isobaths_50mto1500m.shp")
iso_sf_utm <- st_transform(iso_sf, crs = 32610)

# isobath 1500 m
iso_sf1500 <- st_read("../../Data/Environment/iso_sf1500.shp")

# Extent of study area for predictions (OR coastal waters only)
extent_study_area <-  extent(-125.32, -123, 42, 46.3)
polygon_study_area <- as(extent_study_area, 'SpatialPolygons') %>%
  st_as_sf() %>%
  st_set_crs(value = 4326)
polygon_study_area_utm <- polygon_study_area %>% st_transform(crs = 32610)

# bathy contours as provided by NOAA
#https://www.fisheries.noaa.gov/inport/item/54364
extent_study_area_wide <-  extent(-129, -122.3, 37.5, 48) # min and max long correspond to min and max of krill_day_df

# Habitat shapefiles
load("../3-Krill_SDM/Outputs/zones_split.RData")
```

Bongo nets
```{r}
load("../2-Environment/Outputs/biomass_df.RData")
nrow(biomass_df)
biomass_df <- biomass_df %>% 
  # remove daytime bongos except for 092118NH65 because it was more or less dusk
  filter(Day_Night != "D" | Sample.Code == "092118NH65") %>% 
  # change names of columns and compute density and mass by species
  rename(mass_EPAC_juv = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Juvenile",
         mass_EPAC_adu = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Adult",
         mass_EPAC_adf = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Female..Adult",
         mass_EPAC_adm = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Male..Adult",
         
         mass_TSPIN_juv = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Juvenile",
         mass_TSPIN_adu = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Adult",
         mass_TSPIN_adf = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Female..Adult",
         mass_TSPIN_adm = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Male..Adult",
         
         dens_EPAC_juv = "Density....m3..EUPHAUSIA.PACIFICA_Juvenile",
         dens_EPAC_adu = "Density....m3..EUPHAUSIA.PACIFICA_Adult",
         dens_EPAC_adf = "Density....m3..EUPHAUSIA.PACIFICA_Female..Adult",
         dens_EPAC_adm = "Density....m3..EUPHAUSIA.PACIFICA_Male..Adult",
         
         dens_TSPIN_juv = "Density....m3..THYSANOESSA.SPINIFERA_Juvenile",
         dens_TSPIN_adu = "Density....m3..THYSANOESSA.SPINIFERA_Adult",
         dens_TSPIN_adf = "Density....m3..THYSANOESSA.SPINIFERA_Female..Adult",
         dens_TSPIN_adm = "Density....m3..THYSANOESSA.SPINIFERA_Male..Adult") %>% 
  mutate(mass_EPAC = mass_EPAC_juv + mass_EPAC_adu + mass_EPAC_adf + mass_EPAC_adm,
         mass_TSPIN = mass_TSPIN_juv + mass_TSPIN_adu + mass_TSPIN_adf + mass_TSPIN_adm,
         mass_tot = mass_EPAC + mass_TSPIN,
         dens_tot = dens_EPAC_adu + dens_EPAC_juv + dens_EPAC_adf + dens_EPAC_adm +
                    dens_TSPIN_adu + dens_TSPIN_juv + dens_TSPIN_adf + dens_TSPIN_adm) %>% 
  st_as_sf() %>% 
  st_join(zones_split) %>% 
  # remove geometry because it makes the brt / gam loop crash
  as.data.frame() %>% 
  select(-geometry) %>% 
  # add log depth
  mutate(logDEPTH = log(-DEPTH))
nrow(biomass_df)
save(biomass_df, file = "./Outputs/biomass_df.RData")

### biomass NH Line
load("../2-Environment/Outputs/biomass_nhl_df.RData")

biomass_nhl_df <- biomass_nhl_df %>% 
  rename(mass_EPAC = EUPHAUSIA.PACIFICA, mass_TSPIN = THYSANOESSA.SPINIFERA) %>% 
  mutate(mass_tot = mass_EPAC + mass_TSPIN)
save(biomass_nhl_df, file = "./Outputs/biomass_nhl_df.RData")
```

Functions
```{r}
# create custom function to run BRTs in a hierarchical approach
Fun_mybrt <- function(d, v, fam){
  # find what columns the names of the variables correspond to
  var_num <- match(v, names(d))
  var_y <- match("mass", names(d))
  # set starting learning rate (usually between 0.01 and 0.001)
  lr <- 0.001 # this was 0.01 in the  nasc version but here with biomass with 0.01 I tended to have target.trees < 1000 for several rounds of the loop

  target.trees <- 0 # resets the target trees , make sure set condition to lower than 1000

  while (target.trees < 1000) {
  
    mod.brt <- gbm.step(d,
                     gbm.x = var_num, # Your predictor variables, referring to columns in your data frame
                     gbm.y = var_y, # Your response variable
                     family = fam,
                     n.trees = 50,
                     bag.fraction = 0.8, # 0.75 is typical for larger dataset, but can increase bag fraction if dataset is small 
                     plot.main = F,   
                     learning.rate = lr,
                     tree.complexity = 2,
                     verbose = F)  
  
    if (object.size(mod.brt) > 0) {    # > 0 = good, check if hasn't crashed
      target.trees <- mod.brt$gbm.call$best.trees
    }
    else {
      target.trees <- 0  #check if model size is zero 
    }
    lr <- lr / 2
  }
  out <- tibble(model = list(mod.brt),
                lr = lr)
  return(out)
}

# create function to make dataframe predictions from a hierarchical modelling approach: binomial and gamma models
# to specific periods stored in env layers ras
# m_name is the model name = predictors chosen
# here fam should be a vector of the two families bin and logmass
# method is to indicate if the model is a GAM or BRT
Fun_Hpredras <- function(m_name, fam, m_tb, ras, met){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep <- tb %>% 
      filter(model_name == m_name & family == fam & method == met) %>% 
      pull(model_var)
    # model of presence
    m1 <- tb %>% 
      filter(model_name == m_name & family == fam[1] & method == met) %>% 
      pull(model)
    load(m1[[1]])
    if(met == "BRT"){m1m <- mod_brt$model[[1]]}
    if(met == "GAM"){m1m <- mod_gam}
    # model of abundance
    m2 <- tb %>% 
      filter(model_name == m_name & family == fam[2] & method == met) %>% 
      pull(model)
    load(m2[[1]])
    if(met == "BRT"){m2m <- mod_brt$model[[1]]}
    if(met == "GAM"){m2m <- mod_gam}
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week <- r_week[[var_tokeep[[1]]]]
            # then make predictions on raster!
            if(met == "BRT"){
              pred1 <- raster::predict(r_week, m1m, type = "response",
                                      n.trees = m1m$gbm.call$best.trees)
              pred2 <-  raster::predict(r_week, m2m, type = "response",
                                      n.trees = m2m$gbm.call$best.trees)
              pred2 <- exp(pred2)} # response is logged in brt
            if(met == "GAM"){
              pred1 <- raster::predict(r_week, m1m, type = "response")
              pred2 <- raster::predict(r_week, m2m, type = "response")
            }
            predH <- pred1 * pred2
            return(predH)
          }))
          # average across four weeks with NA.rm
          mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
          return(mean_pred)
      }))
  })
}

## BRT bin x GAM abun
Fun_crossENSpredras <- function(m_name, m_tb, ras){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep_bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "GAM") %>% 
      pull(model_var)
    var_tokeep_abu <- tb %>% 
      filter(model_name == m_name & family == "logmass" & method == "GAM") %>% 
      pull(model_var)
    # model of presence
    m.bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "BRT") %>% 
      pull(model)
    load(m.bin[[1]])
    m.bin.m <- mod_brt$model[[1]]
    # model of abundance
    m.abu <- tb %>% 
      filter(model_name == m_name & family == "logmass" & method == "GAM") %>% 
      pull(model)
    load(m.abu[[1]])
    m.abu.m <- mod_gam
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week_bin <- r_week[[var_tokeep_bin[[1]]]]
            r_week_abu <- r_week[[var_tokeep_abu[[1]]]]
            # then make predictions on raster!
            pred1 <- raster::predict(r_week_bin, m.bin.m, type = "response",
                                      n.trees = m.bin.m$gbm.call$best.trees)
            pred2 <- raster::predict(r_week_abu, m.abu.m, type = "response")
            predH <- pred1 * pred2
            return(predH)
        }))
        # average across four weeks with NA.rm
        mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
        return(mean_pred)
      }))
  })
}

# create function to turn tibble of rasters into one dataframe that can generate a ggplot
Fun_meanpredmaps <- function(ras_tb){
  
  # transform into sf format
  d <- ras_tb %>% 
    mutate(pred_df = map(meanpred, function(r){
      # turn raster to sf object
      d <- r %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
      # add coordinates to dataframe of predictions
      d[c("x", "y")] <- st_coordinates(d)
      return(d)
    })) %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = year(as.POSIXct(pred_months, format = "X%Y.%m.%d")),
           period = paste(month, year)) %>%  
    dplyr::select(period, month, year, pred_df) %>%
    # arrange and mutate factor levels so that plots show up in the right order by months 
    arrange(year, month) %>% 
    mutate(period = factor(period, levels = unique(period))) %>% 
    unnest(pred_df)
  
  # create facetted maps by year in columns and months in rows
  g <- ggplot() +
    geom_tile(data = d, aes(x, y, fill = log(fit + 0.1))) +
    scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog biomass",
                         n.breaks = 6) +  
    geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
    geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
    facet_wrap(~period, ncol = n_distinct(d$year), dir = "v") +
    coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
    mon_theme +
    xlab("") +
    ylab("") +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
    
    return(g)
}

# a derivative of the function meanpredmaps but applied to biomass proportions by species
Fun_propmaps <- function(ras_tb){
  
  # transform into sf format
  d <- ras_tb %>% 
    mutate(pred_df = map(meanpred, function(r){
      # turn raster to sf object
      d <- r %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
      # add coordinates to dataframe of predictions
      d[c("x", "y")] <- st_coordinates(d)
      return(d)
    })) %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = year(as.POSIXct(pred_months, format = "X%Y.%m.%d")),
           period = paste(month, year)) %>%  
    dplyr::select(period, month, year, pred_df) %>%
    # arrange and mutate factor levels so that plots show up in the right order by months 
    arrange(year, month) %>% 
    mutate(period = factor(period, levels = unique(period))) %>% 
    unnest(pred_df)
  
  # create facetted maps by year in columns and months in rows
  g <- ggplot() +
    geom_tile(data = d, aes(x, y, fill = fit*100)) +
    scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \n Proportion (%)",
                         n.breaks = 6) +  
    geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
    geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
    facet_wrap(~period, ncol = n_distinct(d$year), dir = "v") +
    coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
    mon_theme +
    xlab("") +
    ylab("") +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
    
    return(g)
}

# create facetted average 2018-2022 maps by months for GAM and BRT
Fun_YearAvgMap <- function(dd, method, group_name, model_name){
    g <- ggplot() +
      geom_tile(data = dd, aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog biomass",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~month, ncol = 6, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
      ggtitle(paste(method, group_name))
    ggsave(g, file = paste0("./Outputs/", method, "_", model_name, "_YearRound_2018-2022_", group_name,".png"), width = 230, height = 150, dpi = 500, units = "mm")
}

# create facetted average 2018-2022 maps by months for GAM and BRT with Blue whales!!
Fun_YearAvgMapBL <- function(dd, method, group_name){
    g <- ggplot() +
      geom_tile(data = dd, aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog biomass",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~month, ncol = 6, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
      ggtitle(paste(method, group_name)) +
      geom_sf(data = bl_df, col = "black", shape = 21, fill = "white", alpha = 0.5)
    ggsave(g, file = paste0("./Outputs/", method, "_dynlogdep_YearRoundBL_2018-2022_", group_name,".png"), width = 230, height = 150, dpi = 500, units = "mm")
}
```


# Running hierarchical SDMs
## Set up models

### Predictor formulas
```{r}
form_chr <- list(
     "topographic" = " mass ~
                   s(DEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5)",
     "topographic_logdep" = " mass ~
                   s(logDEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5)",
     "dynamic" = "mass ~ 
                   s(DEPTH, k = 5) + 
                   s(DIS_CANYON, k = 3) + 
                   s(SLOPE, k = 3) + 
                   s(sst, k = 3) + 
                   s(ssh, k = 3) + 
                   s(sshsd, k = 3) + 
                   s(sstsd, k = 3) + 
                   s(eke, k = 3) +
                   s(curl, k = 3) + 
                   s(ild_05, k = 3) + 
                   s(bbv_200, k = 3)",
     "dynamic_logdep" = "mass ~ 
                   s(logDEPTH, k = 5) +
                   s(DIS_CANYON, k = 3) + 
                   s(SLOPE, k = 3) + 
                   s(sst, k = 3) + 
                   s(ssh, k = 3) + 
                   s(sshsd, k = 3) + 
                   s(sstsd, k = 3) + 
                   s(eke, k = 3) +
                   s(curl, k = 3) + 
                   s(ild_05, k = 3) + 
                   s(bbv_200, k = 3)",
     "dynamic_chla" = "mass ~ 
                   s(DEPTH, k = 5) + 
                   s(DIS_CANYON, k = 3) + 
                   s(SLOPE, k = 3) + 
                   s(sst, k = 3) + 
                   s(ssh, k = 3) + 
                   s(sshsd, k = 3) + 
                   s(sstsd, k = 3) + 
                   s(eke, k = 3) +
                   s(curl, k = 3) + 
                   s(ild_05, k = 3) + 
                   s(bbv_200, k = 3) +
                   s(chla, k = 3)",
      "dynamic_0.6cor" = "mass ~ 
                   s(DEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(sst, k = 5) + 
                   s(sstsd, k = 5) + 
                   s(sshsd, k = 5) + 
                   s(ild_05, k = 5) + 
                   s(bbv_200, k = 5)",
     "dynamic_sst" = "mass ~
                   s(logDEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5) + 
                   s(sst, k = 5)")

model_type <- c(rep("topo", 2), rep("dyn", 5))
model_var <- list(
                   c("DEPTH","SLOPE", "DIS_CANYON"),
                   c("logDEPTH","SLOPE", "DIS_CANYON"),
                   c("DEPTH","DIS_CANYON", "SLOPE", "sst", "ssh", "sshsd", "sstsd","eke", "curl", "ild_05", "bbv_200"),
                   c("logDEPTH","DIS_CANYON", "SLOPE", "sst", "ssh", "sshsd", "sstsd","eke", "curl", "ild_05", "bbv_200"),
                   c("DEPTH","DIS_CANYON", "SLOPE", "sst", "ssh", "sshsd", "sstsd","eke", "curl", "ild_05", "bbv_200", "chla"),
                   c("DEPTH", "SLOPE","sst", "sshsd", "sstsd", "ild_05", "bbv_200"),
                   c("logDEPTH","SLOPE", "DIS_CANYON", "sst"))
```

### Folds and model types
```{r}
# because there are at least two cruises in each season (feb, may, sep) we can leave one out at each run and still have the three seasons represented in the training dataset
Fun_folding <- function(df){
  # set seed to NULL so that we can rerun multiple times until we get a combination of folds without zero presence
  folds <- splitTools::create_folds(df$CruiseID, k = 10, seed = 1, type = "grouped")
  d_cv <- llply(folds, function(f){
    df$fold <- "test"
    df[f,]$fold <- "train"
    return(df)
  })
  return(d_cv)
}

unique(biomass_df$CruiseID)

# there are ten folds in the bongo biomass dataset (vs 8 in the nasc) because the 2018 february/march cruise does not have nasc data
krillmass_cv <- tibble(folds = paste0("Fold", c(1:10)),
                   data = Fun_folding(biomass_df))

modelsH_krillmass_tb <- tibble(model_form = form_chr) %>%
    mutate(model_name = names(form_chr),
         model_type = model_type,
         model_var = model_var,
         model_form = map(model_form, as.formula),
         data = rep(list(krillmass_cv), length(form_chr))) %>% 
  unnest(cols = c(data))
modelsH_krillmass_tb

# concatenate the tibble 2 times for the two krill species
modelsH_krillmass_tb <- rbind(modelsH_krillmass_tb, modelsH_krillmass_tb) %>% 
  mutate(group = rep(c("EPAC", "TSPIN"), each = nrow(modelsH_krillmass_tb))) %>% 
  # create a mass variable by species
  mutate(data = pmap(.l = list(x = data, y = group), .f = function(x, y){
    # add the new variable by sp
    x$mass <- as.data.frame(x[,paste0("mass_", y)])[, 1]
    return(x)
  }))
```

### Run models in parallel
```{r, eval = F}
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  ## BINOMIAL MODELS
  mutate(bin = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, gr = group, fo = folds), .f = function(a, b, c, mn, gr, fo){ 
      # create data set with binomial response
      a_modified <- a %>% mutate(mass = ifelse(mass == 0, 0, 1))
      # in the case of GAMs, the explanatory variables should not have NAs
      a_modified_gam <- a_modified[complete.cases(a_modified[b]), ]
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified, fold == "train"), 
                           v = b, fam = "bernoulli")
      mod_gam <- gam(c, family = binomial(link = "logit"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_bin_", gr, "_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_bin_", gr, "_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return models in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    })) %>%
  ## POSITIVE VALUES MODEL
  mutate(logmass = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, gr = group, fo = folds), .f = function(a, b, c, mn, gr, fo){ 
      # create data set with positive biomass response
      a_modified_gam <- a %>% 
                    filter(mass > 0)
      a_modified_gam <- a_modified_gam[complete.cases(a_modified_gam[b]), ]
      # in the case of BRT the log should be applied prior to running the model
      a_modified_brt <- a %>% 
                    filter(mass > 0) %>% 
                    mutate(mass = log(mass)) # apply a log link
      
      print(paste0("m_abu_", gr, "_", mn, "_", fo))    
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified_brt, fold == "train"), 
                           v = b, fam = "gaussian")
      mod_gam <- gam(c, family = Gamma(link = "log"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_logmass_", gr, "_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_logmass_", gr, "_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return path to model objects in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    
    })) 

# pivot tibble and calculate deviance explained by each model
# in training and in test
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  pivot_longer(cols = c(bin, logmass), names_to = "family", values_to = "mod") %>% 
  unnest(cols = c(mod))

# correct data depending on family and method
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  mutate(data = pmap(list(d = data, v = model_var, fa = family, m = method), 
                     function(d, v, fa, m){
      if(fa == "bin" & m == "BRT"){
        dd <- d %>% mutate(mass = ifelse(mass == 0, 0, 1))
      }
      if(fa == "logmass" & m == "BRT"){
        dd <- d %>% filter(mass > 0) %>% mutate(mass = log(mass))
      }
      if(fa == "bin" & m == "GAM"){
        dd <- d %>% mutate(mass = ifelse(mass == 0, 0, 1))
        if(!is.null(v)){dd <- dd[complete.cases(dd[v]), ]}
      }
      if(fa == "logmass" & m == "GAM"){
        dd <- d %>% filter(mass > 0)
      }
      return(dd)
  }))

save(modelsH_krillmass_tb, file = "./Outputs/modelsH_krillmass_tb.RData")
```

# Model performance

## Performance by method by family
```{r, warning = F, message = F, eval = F}
# this is big!
load("./Outputs/modelsH_krillmass_tb.RData")

modelsH_krillmass_perf <- modelsH_krillmass_tb %>%
  mutate(dev.exp = pmap_dbl(list(mo = model, me = method), function(mo, me){
    load(mo)
    if(me == "BRT"){
      dev <- round((mod_brt$model[[1]]$self.statistics$mean.null - mod_brt$model[[1]]$cv.statistics$deviance.mean)/mod_brt$model[[1]]$self.statistics$mean.null*100, 1)
    }
    if(me == "GAM"){
      dev <- round(summary(mod_gam)$dev.expl*100, 1)
    }
    rm(mo)
    return(dev)
  })) %>%
  # AUC calculation does not depend on method, but can only be done for the bin models
  mutate(AUC.train = pmap_dbl(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    pred.train <- predict(m, subset(d, fold == "train"), type = "response")
    if(fa == "bin"){
        auc <- round(pROC::roc(subset(d, fold == "train")$mass, pred.train)$auc, 3)
    } else {auc <- NA}
    return(auc)
  }))

# that's for testing stuff!
#d <- modelsH_krillmass_perf %>%  filter(folds == "Fold2", model_name == "dynamic_chla", family == "logmass", method == "GAM", group == "TSPIN")
#d <- d$data[[1]]
#mo <- "./Outputs/models/m_gam_logmass_TSPIN_dynamic_chla_Fold2.RData"
#me = "GAM"
#fa = "logmass"


modelsH_krillmass_perf <- modelsH_krillmass_perf %>%
  # For the test fold, we calculate either AUC for the bin model, or pear.cor/spea.cor for the logmass model
  mutate(metrics.test = pmap(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    print(mo)
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    test.data <- subset(d, fold == "test")
    test.data$pred <- predict(m, test.data, type = "response")
    if(fa == "bin"){ # binomial models need zeros and ones
      if(nrow(subset(test.data, !is.na(pred) & mass > 0)) > 1 &
        nrow(subset(test.data, !is.na(pred) & mass == 0)) > 1){
          auc <- round(pROC::roc(test.data$mass, test.data$pred)$auc, 3)
          spea.cor <- NA
          pear.cor <- NA
      } else { spea.cor <- NA ; auc <- NA ; pear.cor <- NA}
    } else { # abundance models (no zeros)
      if(nrow(subset(test.data, !is.na(pred))) > 2){
          auc <- NA
          spea.cor <- round(cor.test(test.data$mass, test.data$pred, method = "spearman")$estimate, 3)
          pear.cor <- round(cor.test(test.data$mass, test.data$pred, method = "pearson")$estimate, 3)
      } else { spea.cor <- NA ; auc <- NA ; pear.cor <- NA}
    }
    metrics <- tibble(spea.cor = spea.cor, pear.cor = pear.cor, AUC.test = auc)
    return(metrics)
  })) %>% 
  unnest(cols = c(metrics.test)) %>% 
  dplyr::select(group, model_name, family, method, folds, dev.exp, AUC.train, AUC.test, pear.cor, spea.cor) %>% 
  arrange(group, family)

save(modelsH_krillmass_perf, file = "./Outputs/modelsH_krillmass_perf.RData")
```


```{r, warning = F, message = F}
load("./Outputs/modelsH_krillmass_perf.RData")

(mean_perf <- modelsH_krillmass_perf %>% 
  group_by(group, model_name, family, method) %>% 
  summarize(mean.dev.exp = mean(dev.exp, na.rm = T),
            mean.AUC.train =  mean(AUC.train, na.rm = T),
            mean.AUC.test = mean(AUC.test, na.rm = T),
            mean.pear.cor = mean(pear.cor, na.rm = T),
            mean.spea.cor = mean(spea.cor, na.rm = T)) %>% 
  mutate_if(is.numeric, ~ round(., digits = 3)) %>% 
  select(group, family, everything()) %>% 
  arrange(group, family, -mean.dev.exp))

openxlsx::write.xlsx(mean_perf, file = "./Outputs/mean_perf.xlsx", overwrite = T)

```

Take a closer look at biomass of tspin because deviance is very low!
```{r, eval = F, echo = F}
d <- modelsH_krillmass_tb %>%  filter(folds == "Fold1", model_name == "dynamic", family == "logmass", method == "GAM", group == "TSPIN")
dabu <- d$data[[1]]
d <- modelsH_krillmass_tb %>%  filter(folds == "Fold1", model_name == "dynamic", family == "bin", method == "GAM", group == "TSPIN")
dbin <- d$data[[1]]

table(dbin$mass)
hist(dabu$mass)

mo <- modelsH_krillmass_tb %>%  filter(folds == "Fold2", model_name == "dynamic_logdep", family == "bin", method == "BRT", group == "TSPIN") %>% pull(model)
load(mo[[1]])
mod_brt
m <- mod_brt$model[[1]]
summary(m)
print(m)
gbm.plot.fits(m)
find.int <- gbm.interactions(m)
find.int$interactions
find.int$rank.list
gbm.perspec(m, 5, 1)
(m$self.statistics$mean.null - m$cv.statistics$deviance.mean)/m$self.statistics$mean.null*100

m$self.statistics$null
m$cv.statistics$deviance.se
mean(m$residuals^2)

ggplot(dbin, aes(logDEPTH, mass)) + geom_point()
ggplot(dbin, aes(ssh, mass)) + geom_point()

residuals(m)

## looks at results from all folds
modelsH_krillmass_perf %>% 
  filter(model_name == "dynamic_logdep", 
         family == "bin", method == "BRT", group == "TSPIN")

biomass_df %>% 
  mutate(tspin_bin = ifelse(mass_TSPIN > 0, 1, 0)) %>% 
  ggplot(., aes(utmx, utmy, col = as.factor(tspin_bin))) +
    geom_point(size = 2) +
    coord_fixed()

a <- biomass_df %>% 
  mutate(tspin_bin = ifelse(mass_TSPIN > 0, 1, 0))
table(a$tspin_bin)

biomass_df %>% 
  filter(utmx < 2.5e+05 & mass_TSPIN > 0)

modelsH_krillmass_perf %>% 
  filter(model_name == "dynamic_logdep", 
         family == "bin", method == "BRT", group == "EPAC")
```


```{r, warning = F, message = F}
g <- ggplot(mean_perf, aes(x = model_name, y = mean.dev.exp, fill = method)) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_grid(family~group) +
    ylab("Mean deviance explained (10 folds CV) in model of krill presence") +
    mon_theme +
    theme(axis.text.x = element_text(angle = 30, hjust = 0.8))
g
ggsave(g, file = "./Outputs/modelH_krillmass_dev.exp.png", width = 250, height = 130, units = "mm", dpi = 600)

g1 <- ggplot(mean_perf, aes(x = model_name, y = mean.AUC.train, fill = method)) +
    geom_bar(stat = "identity", position = "dodge", show.legend = F) +
    ylab("Mean AUC on train set (10 folds CV)") +
  facet_wrap(~group)
g2 <- ggplot(mean_perf, aes(x = model_name, y = mean.AUC.test, fill = method)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab("Mean AUC on test set (10 folds CV) in model of krill presence")+
  facet_wrap(~group)
g <- (g1 + g2) & mon_theme &
    theme(axis.text.x = element_text(angle = 30, hjust = 0.8))
g
ggsave(g, file = "./Outputs/modelH_krillmass_auc.png", width = 250, height = 100, units = "mm", dpi = 600)

g <- ggplot(subset(mean_perf, family == "logmass"), aes(x = model_name, y = mean.pear.cor, fill = method)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab("Mean Pearson cor (10 folds CV) in model of krill abundance conditional on presence") +
    mon_theme +
    facet_wrap(~group) +
    theme(axis.text.x = element_text(angle = 30, hjust = 0.8))
g
ggsave(g, file = "./Outputs/modelH_krillmass_pear.cor.png", width = 250, height = 150, units = "mm", dpi = 600)


g <- ggplot(subset(mean_perf, family == "logmass"), aes(x = model_name, y = mean.spea.cor, fill = method)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab("Mean spearman cor (10 folds CV) in model of krill abundance conditional on presence") +
    mon_theme +
    facet_wrap(~group) +
    theme(axis.text.x = element_text(angle = 30, hjust = 0.8))
g
ggsave(g, file = "./Outputs/modelH_krillmass_spea.cor.png", width = 250, height = 150, units = "mm", dpi = 600)
```

# Predicted maps

## Topographic model

```{r, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")
     
modelsH_krillmass_topo <- modelsH_krillmass_tb %>% 
  filter(model_name == "topographic")

save(modelsH_krillmass_topo, file = "./Outputs/modelsH_krillmass_topo.RData")
```

Predict over all months
```{r}
# select combination of months (May, July, September) and years 2016 onwards
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full prediction rasters                   
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)
```

### BRT

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_topo.RData")

## BRT
predyear_brtH_topo_mass <- ddply(modelsH_krillmass_topo, .(group), function(d) {
  Fun_Hpredras(m_name = "topographic", fam = c("bin","logmass"), m_tb = d, ras = env_pred_yearround, met = "BRT")})
save(predyear_brtH_topo_mass, file = "./Outputs/predyear_brtH_topo_mass.RData")

predyear_brtHavg_topo_mass <-  ddply(predyear_brtH_topo_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_brtHavg_topo_mass, file = "./Outputs/predyear_brtHavg_topo_mass.RData")
```

Average by month
```{r, message = F}
load("./Outputs/predyear_brtHavg_topo_mass.RData")

# average by rasters by month and transform into sf format
predyear_brtH2_topo_mass <- predyear_brtHavg_topo_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_brtH2_topo_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "BRT", model_name = "topo", group_name = unique(.y)))
```

### ENS

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_topo.RData")

## BRT
predyear_crossENS_topo_mass <- ddply(modelsH_krillmass_topo, .(group), function(d) {
  Fun_crossENSpredras(m_name = "topographic",  m_tb = d, ras = env_pred_yearround)})
save(predyear_crossENS_topo_mass, file = "./Outputs/predyear_crossENS_topo_mass.RData")

predyear_crossENSavg_topo_mass <- ddply(predyear_crossENS_topo_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_crossENSavg_topo_mass, file = "./Outputs/predyear_crossENSavg_topo_mass.RData")
```

Average by month
```{r, message = F}
load("./Outputs/predyear_crossENSavg_topo_mass.RData")

# average by rasters by month and transform into sf format
predyear_crossENS2_topo_mass <- predyear_crossENSavg_topo_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_crossENS2_topo_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "ENS", model_name = "topo", group_name = unique(.y)))
```


## Topographic logdep model

```{r, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")
     
modelsH_krillmass_topolog <- modelsH_krillmass_tb %>% 
  filter(model_name == "topographic_logdep")

save(modelsH_krillmass_topolog, file = "./Outputs/modelsH_krillmass_topolog.RData")
```

Predict over all months
```{r}
# select combination of months (May, July, September) and years 2016 onwards
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full prediction rasters                   
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)
```

### BRT

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_topolog.RData")

## BRT
predyear_brtH_topolog_mass <- ddply(modelsH_krillmass_topolog, .(group), function(d) {
  Fun_Hpredras(m_name = "topographic_logdep", fam = c("bin","logmass"), m_tb = d, ras = env_pred_yearround, met = "BRT")})
save(predyear_brtH_topolog_mass, file = "./Outputs/predyear_brtH_topolog_mass.RData")

predyear_brtHavg_topolog_mass <-  ddply(predyear_brtH_topolog_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_brtHavg_topolog_mass, file = "./Outputs/predyear_brtHavg_topolog_mass.RData")
```

Average by month
```{r, message = F}
load("./Outputs/predyear_brtHavg_topolog_mass.RData")

# average by rasters by month and transform into sf format
predyear_brtH2_topolog_mass <- predyear_brtHavg_topolog_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_brtH2_topolog_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "BRT", model_name = "topo_logdep", group_name = unique(.y)))
```

### ENS

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_topolog.RData")

## BRT
predyear_crossENS_topolog_mass <- ddply(modelsH_krillmass_topolog, .(group), function(d) {
  Fun_crossENSpredras(m_name = "topographic_logdep",  m_tb = d, ras = env_pred_yearround)})
save(predyear_crossENS_topolog_mass, file = "./Outputs/predyear_crossENS_topolog_mass.RData")

predyear_crossENSavg_topolog_mass <- ddply(predyear_crossENS_topolog_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_crossENSavg_topolog_mass, file = "./Outputs/predyear_crossENSavg_topolog_mass.RData")
```

Average by month
```{r, message = F}
load("./Outputs/predyear_crossENSavg_topolog_mass.RData")

# average by rasters by month and transform into sf format
predyear_crossENS2_topolog_mass <- predyear_crossENSavg_topolog_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_crossENS2_topolog_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "ENS", model_name = "topo_logdep", group_name = unique(.y)))
```


## Dynamic sst model

```{r, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")
     
modelsH_krillmass_dynsst <- modelsH_krillmass_tb %>% 
  filter(model_name == "dynamic_sst")

save(modelsH_krillmass_dynsst, file = "./Outputs/modelsH_krillmass_dynsst.RData")
```

Predict over all months
```{r}
# select combination of months (May, July, September) and years 2016 onwards
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full prediction rasters                   
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)
```

### BRT

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_dynsst.RData")

## BRT
predyear_brtH_dynsst_mass <- ddply(modelsH_krillmass_dynsst, .(group), function(d) {
  Fun_Hpredras(m_name = "dynamic_sst", fam = c("bin","logmass"), m_tb = d, ras = env_pred_yearround, met = "BRT")})
save(predyear_brtH_dynsst_mass, file = "./Outputs/predyear_brtH_dynsst_mass.RData")

predyear_brtHavg_dynsst_mass <-  ddply(predyear_brtH_dynsst_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_brtHavg_dynsst_mass, file = "./Outputs/predyear_brtHavg_dynsst_mass.RData")
```

Average by month
```{r, message = F}
load("./Outputs/predyear_brtHavg_dynsst_mass.RData")

# average by rasters by month and transform into sf format
predyear_brtH2_dynsst_mass <- predyear_brtHavg_dynsst_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_brtH2_dynsst_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "BRT", model_name = "dynamic_sst", group_name = unique(.y)))
```

## Dynamic_logdep model
Using the best model on average
```{r, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")
     
modelsH_krillmass_dynlogdep <- modelsH_krillmass_tb %>% 
  filter(model_name == "dynamic_logdep")

save(modelsH_krillmass_dynlogdep, file = "./Outputs/modelsH_krillmass_dynlogdep.RData")
```

### Feb, May, September

```{r}
#Periods of predictions over three months of interest
# select combination of months (Feb, May, September) and years 2018 onwards
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(2, 5, 9))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full prediction rasters                   
env_pred_subset <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)
```

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_dynlogdep.RData")

## Run predictions with a cross-Ensemble combining the BRT binomial model with the GAM abundance model
pred_crossENS_dynlogdep <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_crossENSpredras(m_name = "dynamic_logdep",  m_tb = d, ras = env_pred_subset)})
save(pred_crossENS_dynlogdep, file = "./Outputs/pred_crossENS_dynlogdep_mass.RData")

pred_crossENSavg_dynlogdep <- ddply(pred_crossENS_dynlogdep, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(pred_crossENSavg_dynlogdep, file = "./Outputs/pred_crossENSavg_dynlogdep_mass.RData")
rm(pred_crossENSavg_dynlogdep);rm(pred_crossENS_dynlogdep)

## BRT
pred_brtH_dynlogdep <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_Hpredras(m_name = "dynamic_logdep", fam = c("bin","logmass"), m_tb = d, ras = env_pred_subset, met = "BRT")})
save(pred_brtH_dynlogdep, file = "./Outputs/pred_brtH_dynlogdep_mass.RData")

pred_brtHavg_dynlogdep <- ddply(pred_brtH_dynlogdep, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(pred_brtHavg_dynlogdep, file = "./Outputs/pred_brtHavg_dynlogdep_mass.RData")
rm(pred_brtHavg_dynlogdep);rm(pred_brtH_dynlogdep)

## GAM
pred_gamH_dynlogdep <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_Hpredras(m_name = "dynamic_logdep", fam = c("bin","logmass"), m_tb = d, ras = env_pred_subset, met = "GAM")})
save(pred_gamH_dynlogdep, file = "./Outputs/pred_gamH_dynlogdep_mass.RData")

pred_gamHavg_dynlogdep <- ddply(pred_gamH_dynlogdep, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(pred_gamHavg_dynlogdep, file = "./Outputs/pred_gamHavg_dynlogdep_mass.RData")
rm(pred_gamHavg_dynlogdep);rm(pred_gamH_dynlogdep)
```

Generate maps by month by year by species
```{r, fig.width = 15, fig.height = 11, warning = F}
load("./Outputs/pred_gamHavg_dynlogdep_mass.RData")
load("./Outputs/pred_brtHavg_dynlogdep_mass.RData")
load("./Outputs/pred_crossENSavg_dynlogdep_mass.RData")

d_ply(pred_crossENSavg_dynlogdep, .(group), function(d){
  map <- Fun_meanpredmaps(d)
  map <- map + ggtitle(d$group[[1]])
  ggsave(map, file = paste0("./Outputs/crossENS_dynlogdep_FebMaySep_2018-2022_", d$group[[1]],".png"), width = 180, height = 180, dpi = 500, units = "mm")
  grid.arrange(map)
})

d_ply(pred_gamHavg_dynlogdep, .(group), function(d){
  map <- Fun_meanpredmaps(d)
  map <- map + ggtitle(d$group[[1]])
  ggsave(map, file = paste0("./Outputs/GAMH_dynlogdep_FebMaySep_2018-2022_", d$group[[1]],".png"), width = 180, height = 180, dpi = 500, units = "mm")
  grid.arrange(map)
})

d_ply(pred_brtHavg_dynlogdep, .(group), function(d){
  map <- Fun_meanpredmaps(d)
  map <- map + ggtitle(d$group[[1]])
  ggsave(map, file = paste0("./Outputs/BRTH_dynlogdep_FebMaySep_2018-2022_", d$group[[1]],".png"), width = 180, height = 180, dpi = 500, units = "mm")
  grid.arrange(map)
})
```

### Year-round

Predict over all months
```{r}
# select combination of months (May, July, September) and years 2016 onwards
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full prediction rasters                   
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)
```

Run predictions
```{r, eval = F}
load("./Outputs/modelsH_krillmass_dynlogdep.RData")

## BRT
predyear_brtH_dynlogdep_mass <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_Hpredras(m_name = "dynamic_logdep", fam = c("bin","logmass"), m_tb = d, ras = env_pred_yearround, met = "BRT")})
save(predyear_brtH_dynlogdep_mass, file = "./Outputs/predyear_brtH_dynlogdep_mass.RData")

predyear_brtHavg_dynlogdep_mass <-  ddply(predyear_brtH_dynlogdep_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_brtHavg_dynlogdep_mass, file = "./Outputs/predyear_brtHavg_dynlogdep_mass.RData")

## GAM
predyear_gamH_dynlogdep_mass <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_Hpredras(m_name = "dynamic_logdep", fam = c("bin","logmass"), m_tb = d, ras = env_pred_yearround, met = "GAM")})
save(predyear_gamH_dynlogdep_mass, file = "./Outputs/predyear_gamH_dynlogdep_mass.RData")

predyear_gamHavg_dynlogdep_mass <-  ddply(predyear_gamH_dynlogdep_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_gamHavg_dynlogdep_mass, file = "./Outputs/predyear_gamHavg_dynlogdep_mass.RData")

## ENS
predyear_crossENS_dynlogdep_mass <- ddply(modelsH_krillmass_dynlogdep, .(group), function(d) {
  Fun_crossENSpredras(m_name = "dynamic_logdep",  m_tb = d, ras = env_pred_yearround)})
save(predyear_crossENS_dynlogdep_mass, file = "./Outputs/predyear_crossENS_dynlogdep_mass.RData")

predyear_crossENSavg_dynlogdep_mass <- ddply(predyear_crossENS_dynlogdep_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_crossENSavg_dynlogdep_mass, file = "./Outputs/predyear_crossENSavg_dynlogdep_mass.RData")
```

### Average by month

From the ensemble model
```{r, message = F}
load("./Outputs/predyear_crossENSavg_dynlogdep_mass.RData")

# average by rasters by month and transform into sf format
predyear_crossENS2_dynlogdep_mass <- predyear_crossENSavg_dynlogdep_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })

predyear_crossENS2_dynlogdep_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "ENS", model_name = "dynlogdep", group_name = unique(.y)))
```

From the GAM hierarchical model
```{r, message = F}
load("./Outputs/predyear_gamHavg_dynlogdep_mass.RData")

# average by rasters by month and transform into sf format
predyear_gamH2_dynlogdep_mass <- predyear_gamHavg_dynlogdep_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })

predyear_gamH2_dynlogdep_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "GAM", model_name = "dynlogdep", group_name = unique(.y)))
```

From the BRT hierarchical model
```{r, message = F}
load("./Outputs/predyear_brtHavg_dynlogdep_mass.RData")

# average by rasters by month and transform into sf format
predyear_brtH2_dynlogdep_mass <- predyear_brtHavg_dynlogdep_mass %>% 
    mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE)) %>% 
    ddply(., .(group, month), function(d){
      out <- mean(stack(d$meanpred), na.rm = T) %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
            # add coordinates to dataframe of predictions
      out[c("x", "y")] <- st_coordinates(out)
      return(out)
    })
predyear_brtH2_dynlogdep_mass %>%
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMap(.x, method = "BRT", model_name = "dynlogdep", group_name = unique(.y)))
```

### Compare to blue whales

Compare predictions to blue whale observations
```{r}
load("../../3-ID_count_extract/Outputs/dettot_df.RData")
bl_opal_df <- dettot_df %>% 
  filter(sp_code == "BLWH") %>% 
  mutate(month = lubridate::month(time_pst, label = TRUE, abbr = TRUE),
         year = year(time_pst),
         period = paste(month, year)) %>%  
  dplyr::select(period, month, year, geometry, sp_count)

bl_df <- read.csv("../../../Data/External_validation_data.csv") %>% 
  filter(Species == "BLWH", Year >= 2016) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 32610) %>% 
  mutate(time = ifelse(DateTime_PST != "", substr(DateTime_PST, 1, 10), substr(DateTime_GMT, 1, 10))) %>% 
  mutate(time = as.POSIXct(time, format = "%Y/%m/%d"),
         month = lubridate::month(time, label = TRUE, abbr = TRUE),
         year = year(time),
         period = paste(month, year)) %>%  
  rename(sp_count = Count) %>% 
  dplyr::select(period, month, year, geometry, sp_count) %>% 
  rbind(bl_opal_df) %>% 
  mutate(period = factor(period, levels = unique(period)))

predyear_brtH2_dynlogdep_mass %>%
  filter(group %in% c("EPAC", "TSPIN")) %>% 
  group_by(group) %>% 
  group_map(~ Fun_YearAvgMapBL(.x, method = "BRT", group_name = unique(.y)))
```

### Month of May across species
```{r}
# BRT
g <- predyear_brtH2_dynlogdep_mass %>% 
  filter(month == "May") %>% 
  ggplot(.) +
      geom_tile(aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nbiomass",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~group, ncol = 4, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
g
ggsave(g, file = "./Outputs/BRT_dynlogdepmass_AvgMay_2018-2022.png", width = 120, height = 110, dpi = 500, units = "mm")

# GAM
g <- predyear_gamH2_dynlogdep_mass %>% 
  filter(month == "May") %>% 
  ggplot(.) +
      geom_tile(aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nbiomass",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~group, ncol = 4, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
ggsave(g, file = "./Outputs/GAM_dynlogdepmass_AvgMay_2018-2022.png", width = 120, height = 110, dpi = 500, units = "mm") 

# ENS
g <- predyear_crossENS2_dynlogdep_mass %>% 
  filter(month == "May") %>% 
  ggplot(.) +
      geom_tile(aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nbiomass",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~group, ncol = 4, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
g
ggsave(g, file = "./Outputs/ENS_dynlogdepmass_AvgMay_2018-2022.png", width = 120, height = 110, dpi = 500, units = "mm") 
```

# Biomass trends in time

```{r}
# BRT #################################################
load("./Outputs/predyear_brtHavg_dynlogdep_mass.RData")

#looking into the crazy values
r=predyear_brtHavg_dynlogdep_mass %>% filter(group == "TSPIN", pred_months == "X2018.7.24") %>% pull(meanpred)
r<- r[[1]]
plot(r)
r[r > 500] <- 500
plot(r)

krill_trends_brt <- predyear_brtHavg_dynlogdep_mass %>%
  mutate(sum_mass = map_dbl(meanpred, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(-c(meanpred)) %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

# plots trends by year
g <- ggplot(krill_trends_brt, aes(month, sum_mass, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
g
ggsave(g, file = "./Outputs/BRT_pred_dynlogdep_trends_mass.png", width = 180, height = 70, dpi = 500, units = "mm")

# GAM ##################################################
load("./Outputs/predyear_gamHavg_dynlogdep_mass.RData")

krill_trends_gam <- predyear_gamHavg_dynlogdep_mass %>%
  mutate(sum_mass = map_dbl(meanpred, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(-c(meanpred)) %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

# plots trends by year
g <- ggplot(krill_trends_gam, aes(month, sum_mass, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
g
ggsave(g, file = "./Outputs/GAM_pred_dynlogdep_trends_mass.png", width = 180, height = 70, dpi = 500, units = "mm")
```

# Relationships

## Dynlogdep model
### Variable influence
```{r}
load("./Outputs/modelsH_krillmass_dynlogdep.RData")

## GAM
Fun_VarPVGAM <- function(m){
  load(m)
  data.frame(predvar = names(summary(mod_gam)$chi.sq),
             pv = summary(mod_gam)$s.pv)
}

Fun_VarInfluGAM <- function(d){
  data.frame(count0.001 = length(d$pv[d$pv <= 0.001]),
             count0.01 = length(d$pv[d$pv <= 0.01]),
             count0.05 = length(d$pv[d$pv <= 0.05]),
             mean_pvalue = round(mean(d$pv), 3))
}

influ_gam_dynlogdep_mass <- modelsH_krillmass_dynlogdep %>% 
  filter(method == "GAM") %>% 
  mutate(influence = map(model, function(m) Fun_VarPVGAM(m))) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  ddply(., .(group, method, family, predvar), function(d) Fun_VarInfluGAM(d))
influ_gam_dynlogdep_mass
  
save(influ_gam_dynlogdep_mass, file = "./Outputs/influ_gam_dynlogdep_mass.RData")


### BRT
influ_brt_dynlogdep_mass <- modelsH_krillmass_dynlogdep %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(group, method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)
influ_brt_dynlogdep_mass
save(influ_brt_dynlogdep_mass, file = "./Outputs/influ_brt_dynlogdep_mass.RData")

```

```{r}
influ_gam_plot_mass <- ddply(influ_gam_dynlogdep_mass, .(group, family), function(d){
  d <- d %>%  
    dplyr::select(-c(group, method, family, mean_pvalue)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    dplyr::rename(thresh = predvar, "DEPTH" = "s(logDEPTH)", "CANYON" = "s(DIS_CANYON)", 
           "SLOPE" ="s(SLOPE)", "BBV" = "s(bbv_200)", 
           "CURL" = "s(curl)", "ILD" = "s(ild_05)", "SSHSD" = "s(sshsd)", 
           "SSH" = "s(ssh)", "SSTSD" = "s(sstsd)", "SST" = "s(sst)",
           "EKE" = "s(eke)") %>% 
    select(thresh, DEPTH, CANYON, SLOPE, BBV, ILD, CURL, SST, SSH, SSHSD, SSTSD, EKE)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  d <- cbind(d["thresh"], as.data.frame(a))
  return(d)
})

png("./Outputs/influ_gam_bin_dynlogdep_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_gam_plot_mass %>%
  filter(group == s, family == "bin") %>% 
  column_to_rownames("thresh") %>% 
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(10, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.4),
             caxislabels = c(0, 2, 4, 6, 8, ''),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_gam_abundance_dynlogdep_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_gam_plot_mass %>%
  filter(group == s, family == "logmass") %>% 
  column_to_rownames("thresh") %>% 
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(10, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.4),
             caxislabels = c(0, 2, 4, 6, 8, ''),
             palcex = 0.8)
}
dev.off()
```

```{r}
influ_brt_plot_mass <- ddply(influ_brt_dynlogdep_mass, .(group, family), function(d){
  d <- d %>% 
    ungroup() %>% 
    dplyr::select(-c(group, method, family, sdInf)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    rename("CANYON" = "DIS_CANYON", "BBV" = "bbv_200", 
           "CURL" = "curl", "ILD" = "ild_05", "SSHSD" = "sshsd", 
           "SSH" = "ssh", "SSTSD" = "sstsd", "SST" = "sst", "EKE" = "eke", "DEPTH" = "logDEPTH") %>% 
    select(predvar, DEPTH, CANYON, SLOPE, BBV, ILD, CURL, SST, SSH, SSHSD, SSTSD, EKE)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  return(a)
})

png("./Outputs/influ_brt_bin_dynlogdep_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "bin") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_brt_abundance_dynlogdep_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "logmass") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()
```


## Dynsst model
### Variable influence
```{r}
load("./Outputs/modelsH_krillmass_dynsst.RData")

## GAM
Fun_VarPVGAM <- function(m){
  load(m)
  data.frame(predvar = names(summary(mod_gam)$chi.sq),
             pv = summary(mod_gam)$s.pv)
}

Fun_VarInfluGAM <- function(d){
  data.frame(count0.001 = length(d$pv[d$pv <= 0.001]),
             count0.01 = length(d$pv[d$pv <= 0.01]),
             count0.05 = length(d$pv[d$pv <= 0.05]),
             mean_pvalue = round(mean(d$pv), 3))
}

influ_gam_dynsst_mass <- modelsH_krillmass_dynsst %>% 
  filter(method == "GAM") %>% 
  mutate(influence = map(model, function(m) Fun_VarPVGAM(m))) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  ddply(., .(group, method, family, predvar), function(d) Fun_VarInfluGAM(d))
influ_gam_dynsst_mass
  
save(influ_gam_dynsst_mass, file = "./Outputs/influ_gam_dynsst_mass.RData")


### BRT
influ_brt_dynsst_mass <- modelsH_krillmass_dynsst %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(group, method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)
influ_brt_dynsst_mass
save(influ_brt_dynsst_mass, file = "./Outputs/influ_brt_dynsst_mass.RData")

```

```{r}
influ_gam_plot_mass <- ddply(influ_gam_dynsst_mass, .(group, family), function(d){
  d <- d %>%  
    dplyr::select(-c(group, method, family, mean_pvalue)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    dplyr::rename(thresh = predvar, "DEPTH" = "s(logDEPTH)", "CANYON" = "s(DIS_CANYON)", 
           "SLOPE" ="s(SLOPE)",  "SST" = "s(sst)") %>% 
    select(thresh, DEPTH, CANYON, SLOPE, SST)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  d <- cbind(d["thresh"], as.data.frame(a))
  return(d)
})

png("./Outputs/influ_gam_bin_dynsst_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_gam_plot_mass %>%
  filter(group == s, family == "bin") %>% 
  column_to_rownames("thresh") %>% 
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(10, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.4),
             caxislabels = c(0, 2, 4, 6, 8, ''),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_gam_abundance_dynsst_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_gam_plot_mass %>%
  filter(group == s, family == "logmass") %>% 
  column_to_rownames("thresh") %>% 
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(10, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.4),
             caxislabels = c(0, 2, 4, 6, 8, ''),
             palcex = 0.8)
}
dev.off()
```

```{r}
influ_brt_plot_mass <- ddply(influ_brt_dynsst_mass, .(group, family), function(d){
  d <- d %>% 
    ungroup() %>% 
    dplyr::select(-c(group, method, family, sdInf)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    rename("CANYON" = "DIS_CANYON", "DEPTH" = "logDEPTH", "SST" = "sst") %>% 
    select(predvar, DEPTH, CANYON, SLOPE, SST)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  return(a)
})

png("./Outputs/influ_brt_bin_dynsst_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "bin") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_brt_abundance_dynsst_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "logmass") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()
```


## Topo model
### Variable influence
```{r}
load("./Outputs/modelsH_krillmass_topo.RData")

### BRT
influ_brt_topo_mass <- modelsH_krillmass_topo %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(group, method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)
influ_brt_topo_mass
save(influ_brt_topo_mass, file = "./Outputs/influ_brt_topo_mass.RData")

```

```{r}
influ_brt_plot_mass <- ddply(influ_brt_topo_mass, .(group, family), function(d){
  d <- d %>% 
    ungroup() %>% 
    dplyr::select(-c(group, method, family, sdInf)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    rename("CANYON" = "DIS_CANYON") %>% 
    select(predvar, DEPTH, CANYON, SLOPE)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  return(a)
})

png("./Outputs/influ_brt_bin_topo_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "bin") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_brt_abundance_topo_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "logmass") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()
```

## Topolog model
### Variable influence
```{r}
load("./Outputs/modelsH_krillmass_topolog.RData")

## GAM
Fun_VarPVGAM <- function(m){
  load(m)
  data.frame(predvar = names(summary(mod_gam)$chi.sq),
             pv = summary(mod_gam)$s.pv)
}

Fun_VarInfluGAM <- function(d){
  data.frame(count0.001 = length(d$pv[d$pv <= 0.001]),
             count0.01 = length(d$pv[d$pv <= 0.01]),
             count0.05 = length(d$pv[d$pv <= 0.05]),
             mean_pvalue = round(mean(d$pv), 3))
}

influ_gam_topolog_mass <- modelsH_krillmass_topolog %>% 
  filter(method == "GAM") %>% 
  mutate(influence = map(model, function(m) Fun_VarPVGAM(m))) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  ddply(., .(group, method, family, predvar), function(d) Fun_VarInfluGAM(d))
influ_gam_topolog_mass
  
save(influ_gam_topolog_mass, file = "./Outputs/influ_gam_topolog_mass.RData")

### BRT
influ_brt_topolog_mass <- modelsH_krillmass_topolog %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(group, method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)
influ_brt_topolog_mass
save(influ_brt_topolog_mass, file = "./Outputs/influ_brt_topolog_mass.RData")

```

```{r}
influ_brt_plot_mass <- ddply(influ_brt_topolog_mass, .(group, family), function(d){
  d <- d %>% 
    ungroup() %>% 
    dplyr::select(-c(group, method, family, sdInf)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    rename("CANYON" = "DIS_CANYON") %>% 
    select(predvar, logDEPTH, CANYON, SLOPE)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  return(a)
})

png("./Outputs/influ_brt_bin_topolog_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "bin") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()

png("./Outputs/influ_brt_abundance_topolog_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_brt_plot_mass %>%
  filter(group == s, family == "logmass") %>%
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(25, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.7),
             caxislabels = c(0, 5, 10, 15, 20, 25),
             palcex = 0.8)
}
dev.off()
```

```{r}
influ_gam_plot_mass <- ddply(influ_gam_topolog_mass, .(group, family), function(d){
  d <- d %>%  
    dplyr::select(-c(group, method, family, mean_pvalue)) %>% 
    t %>% # transpose columns
    as.data.frame() %>% 
    rownames_to_column()
  names(d) <- d[1,]
  d <- d[-1,]
  d <- d %>% 
    dplyr::rename(thresh = predvar, "DEPTH" = "s(logDEPTH)", "CANYON" = "s(DIS_CANYON)", 
           "SLOPE" ="s(SLOPE)") %>% 
    select(thresh, DEPTH, CANYON, SLOPE)
  # convert back to numeric
  a <- sapply(d[, names(d)[-1]], as.numeric)
  d <- cbind(d["thresh"], as.data.frame(a))
  return(d)
})

png("./Outputs/influ_gam_abundance_topolog_mass.png", res = 600, units = "mm", width = 80, height = 160)
op <- par(mar = c(0.2, 0.2, 2, 0.2))
par(mfrow = c(2,1))
for(s in unique(krillsp_col$group)){
  d <- influ_gam_plot_mass %>%
  filter(group == s, family == "logmass") %>% 
  column_to_rownames("thresh") %>% 
  dplyr::select(-c(group, family)) %>% 
  dplyr::arrange(desc(row_number()))
  d <- rbind(rep(10, ncol(d)) , rep(0, ncol(d)) , d)
  radarchart(d, seg = 5, axistype=1, 
             pcol = krillsp_col[krillsp_col$group == s, ]$col, 
             plwd=1.5,  plty=1, cglcol="grey", cglty=1,
             axislabcol="grey25", cglwd=0.8, vlcex=1,
             pfcol = scales::alpha(krillsp_col[krillsp_col$group == s, ]$col, 0.4),
             caxislabels = c(0, 2, 4, 6, 8, ''),
             palcex = 0.8)
}
dev.off()
```
## Partial Dependence plots

### Dynlogdep

#### ENS
```{r}
# create a dataframe filled with the average of each variable
# ignore warning
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_dynlogdep$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
ENScurves_krillmass_dynlogdep <- modelsH_krillmass_dynlogdep %>% 
  filter((method == "BRT" & family == "bin") | (method == "GAM" & family == "logmass")) %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    if(me == "BRT"){
      m <- mod_brt$model[[1]]
      var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    }
    if(me == "GAM"){
      m <- mod_gam
      var_selec <- data.frame(predvar = names(summary(m)$chi.sq),
                        pv = summary(m)$s.pv) %>% 
        filter(pv <= 0.05) %>% 
        mutate(var = substr(predvar, 3, (nchar(predvar)-1)))
    }
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        if(me == "BRT"){
          pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
          pred$se.fit <- NA # no error estimate in the BRT
        }
        if(me == "GAM"){
          pred <- dismo::predict(m, mean_pred_x_df, type = "response", se.fit = T)
        }
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

ENScurves_krillmass_dynlogdep <- ENScurves_krillmass_dynlogdep %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(ENScurves_krillmass_dynlogdep, file = "./Outputs/ENScurves_krillmass_dynlogdep.RData")

g1 <- ENScurves_krillmass_dynlogdep %>% 
  filter(family == "bin") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Probability of krill presence (BRT)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g1
ggsave(g1, file = "./Outputs/ENScurves_krillmass_dynlogdep_bin.png", width = 180, height = 120, dpi = 300, units = "mm")

g2 <- ENScurves_krillmass_dynlogdep %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill abundance conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g2
ggsave(g2, file = "./Outputs/ENScurves_krillmass_dynlogdep_abu.png", width = 180, height = 120, dpi = 300, units = "mm")
```

#### BRT
```{r}
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_dynlogdep$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
BRTcurves_krillmass_dynlogdep <- modelsH_krillmass_dynlogdep %>% 
  filter(method == "BRT") %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    m <- mod_brt$model[[1]]
    var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
        pred$se.fit <- NA # no error estimate in the BRT
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

BRTcurves_krillmass_dynlogdep <- BRTcurves_krillmass_dynlogdep %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(BRTcurves_krillmass_dynlogdep, file = "./Outputs/BRTcurves_krillmass_dynlogdep.RData")


g2 <- BRTcurves_krillmass_dynlogdep %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill biomass conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))

ggsave(g2, file = "./Outputs/BRTcurves_krillmass_dynlogdep_abu.png", width = 180, height = 120, dpi = 300, units = "mm")
```

### Dynsst

#### ENS
```{r}
# create a dataframe filled with the average of each variable
# ignore warning
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_dynsst$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
ENScurves_krillmass_dynsst <- modelsH_krillmass_dynsst %>% 
  filter((method == "BRT" & family == "bin") | (method == "GAM" & family == "logmass")) %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    if(me == "BRT"){
      m <- mod_brt$model[[1]]
      var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    }
    if(me == "GAM"){
      m <- mod_gam
      var_selec <- data.frame(predvar = names(summary(m)$chi.sq),
                        pv = summary(m)$s.pv) %>% 
        filter(pv <= 0.05) %>% 
        mutate(var = substr(predvar, 3, (nchar(predvar)-1)))
    }
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        if(me == "BRT"){
          pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
          pred$se.fit <- NA # no error estimate in the BRT
        }
        if(me == "GAM"){
          pred <- dismo::predict(m, mean_pred_x_df, type = "response", se.fit = T)
        }
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

ENScurves_krillmass_dynsst <- ENScurves_krillmass_dynsst %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(ENScurves_krillmass_dynsst, file = "./Outputs/ENScurves_krillmass_dynsst.RData")

g1 <- ENScurves_krillmass_dynsst %>% 
  filter(family == "bin") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Probability of krill presence (BRT)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g1
ggsave(g1, file = "./Outputs/ENScurves_krillmass_dynsst_bin.png", width = 180, height = 120, dpi = 300, units = "mm")

## look at average sst trends
g1sst <- ENScurves_krillmass_dynsst %>% 
  filter(family == "bin", predvar == "sst") %>% 
  mutate(id = paste(folds, group, sep = "_")) %>% 
  group_by(group, x) %>% 
  summarize(meanfit = mean(fit),
            sdfit = sd(fit)) %>% 
  mutate(group = factor(group, levels = c("EPAC","TSPIN"))) %>% 
  ggplot(aes(x, meanfit, col = group)) +
      geom_ribbon(aes(ymin = meanfit - 1.96 * sdfit, ymax = meanfit + 1.96 * sdfit, fill = group), alpha = 0.3) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      ylab("Krill biomass conditional on presence (GAM)") +
      xlab("SST (C)") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))

ggsave(g1sst, file = "./Outputs/ENScurves_krillmass_dynsst_bin_SST.png", width = 100, height = 100, dpi = 300, units = "mm")
g1sst

g2 <- ENScurves_krillmass_dynsst %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill abundance conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g2
ggsave(g2, file = "./Outputs/ENScurves_krillmass_dynsst_abu.png", width = 180, height = 120, dpi = 300, units = "mm")
```

#### BRT
```{r}
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_dynsst$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
BRTcurves_krillmass_dynsst <- modelsH_krillmass_dynsst %>% 
  filter(method == "BRT") %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    m <- mod_brt$model[[1]]
    var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
        pred$se.fit <- NA # no error estimate in the BRT
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

BRTcurves_krillmass_dynsst <- BRTcurves_krillmass_dynsst %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(BRTcurves_krillmass_dynsst, file = "./Outputs/BRTcurves_krillmass_dynsst.RData")


g2 <- BRTcurves_krillmass_dynsst %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE", "BBV", "ILD", "CURL", "SST", "SSH", "SSHSD", "SSTSD", "EKE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill biomass conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))

ggsave(g2, file = "./Outputs/BRTcurves_krillmass_dynsst_abu.png", width = 180, height = 100, dpi = 300, units = "mm")
g2

## look at average sst trends
g2sst <- BRTcurves_krillmass_dynsst %>% 
  filter(family == "logmass", predvar == "sst") %>% 
  mutate(id = paste(folds, group, sep = "_")) %>% 
  group_by(group, x) %>% 
  summarize(meanfit = mean(fit),
            sdfit = sd(fit)) %>% 
  ggplot(aes(x, meanfit, col = group)) +
      geom_ribbon(aes(ymin = meanfit - 1.96 * sdfit, ymax = meanfit + 1.96 * sdfit, fill = group), alpha = 0.3) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      ylab("Krill biomass conditional on presence (GAM)") +
      xlab("SST (C)") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))

ggsave(g2sst, file = "./Outputs/BRTcurves_krillmass_dynsst_abu_SST.png", width = 100, height = 100, dpi = 300, units = "mm")
```

### Topo
#### BRT
```{r}
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_topo$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
BRTcurves_krillmass_topo <- modelsH_krillmass_topo %>% 
  filter(method == "BRT") %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    m <- mod_brt$model[[1]]
    var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
        pred$se.fit <- NA # no error estimate in the BRT
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

BRTcurves_krillmass_topo <- BRTcurves_krillmass_topo %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(BRTcurves_krillmass_topo, file = "./Outputs/BRTcurves_krillmass_topo.RData")

g1 <- BRTcurves_krillmass_topo %>% 
  filter(family == "bin") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Probability of krill presence (BRT)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g1
ggsave(g1, file = "./Outputs/BRTcurves_krillmass_topo_bin.png", width = 180, height = 120, dpi = 300, units = "mm")

g2 <- BRTcurves_krillmass_topo %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON'; 'bbv_200' = 'BBV'; 
           'curl' = 'CURL'; 'ild_05' = 'ILD'; 'sshsd'='SSHSD'; 
           'ssh' = 'SSH';  'sstsd' = 'SSTSD';'sst' = 'SST'; 'eke' = 'EKE';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH", "CANYON", "SLOPE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill biomass conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))

ggsave(g2, file = "./Outputs/BRTcurves_krillmass_topo_abu.png", width = 180, height = 120, dpi = 300, units = "mm")
```

### Topolog
#### ENS
```{r}
load("./Outputs/modelsH_krillmass_topolog.RData")

# create a dataframe filled with the average of each variable
# ignore warning
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_topolog$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
ENScurves_krillmass_topolog <- modelsH_krillmass_topolog %>% 
  filter((method == "BRT" & family == "bin") | (method == "GAM" & family == "logmass")) %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    if(me == "BRT"){
      m <- mod_brt$model[[1]]
      var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    }
    if(me == "GAM"){
      m <- mod_gam
      var_selec <- data.frame(predvar = names(summary(m)$chi.sq),
                        pv = summary(m)$s.pv) %>% 
        filter(pv <= 0.05) %>% 
        mutate(var = substr(predvar, 3, (nchar(predvar)-1)))
    }
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        if(me == "BRT"){
          pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
          pred$se.fit <- NA # no error estimate in the BRT
        }
        if(me == "GAM"){
          pred <- dismo::predict(m, mean_pred_x_df, type = "response", se.fit = T)
        }
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

ENScurves_krillmass_topolog <- ENScurves_krillmass_topolog %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(ENScurves_krillmass_topolog, file = "./Outputs/ENScurves_krillmass_topolog.RData")

g1 <- ENScurves_krillmass_topolog %>% 
  filter(family == "bin") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Probability of krill presence (BRT)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g1
ggsave(g1, file = "./Outputs/ENScurves_krillmass_topolog_bin.png", width = 180, height = 100, dpi = 300, units = "mm")

g2 <- ENScurves_krillmass_topolog %>% 
  filter(family == "logmass") %>% 
  mutate(id = paste(folds, group, sep = "_"),
         predvar = car::recode(predvar, "'DIS_CANYON' = 'CANYON';
           'logDEPTH' = 'DEPTH'", as.numeric = F)) %>% 
  mutate(predvar = factor(predvar, levels = c("DEPTH",  "CANYON", "SLOPE"))) %>% 
  ggplot(aes(x, fit, group = id, col = group)) +
      geom_path(linewidth = 0.8, alpha = 0.9) +
      facet_wrap(~predvar, scales = "free", ncol = 4) + #labeller = labeller(predvar = labels_predvar)
      ylab("Krill abundance conditional on presence (GAM)") +
      xlab("Environmental predictors") +
      mon_theme +
      scale_color_manual(name = "Species & life stage", values = krillsp_pal) +
      scale_fill_manual(name = "Species & life stage", values = krillsp_pal) +
      theme(legend.key.size = unit(1.0, "lines"),
        legend.position = "bottom",
        strip.background = element_rect(color = "black", fill = "white", linewidth = 0.5, linetype = "solid"))
g2
ggsave(g2, file = "./Outputs/ENScurves_krillmass_topolog_abu.png", width = 180, height = 100, dpi = 300, units = "mm")
```

# Pred NASC x Pred Biomass proportion

## Produce biomass proportion maps

Calculate the proportion of EPAC and TSPIN over monthly rasters
```{r}
############ DYNLOGDEP PREDICTORS ##########################
load("./Outputs/predyear_crossENSavg_dynlogdep_mass.RData")
pred_propbiomass_ENSdynlogdep <- ddply(predyear_crossENSavg_dynlogdep_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

load("./Outputs/predyear_brtHavg_dynlogdep_mass.RData")
pred_propbiomass_BRTdynlogdep <- ddply(predyear_brtHavg_dynlogdep_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

load("./Outputs/predyear_gamHavg_dynlogdep_mass.RData")
pred_propbiomass_GAMdynlogdep <- ddply(predyear_gamHavg_dynlogdep_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

############ DYNSST PREDICTORS ##########################
load("./Outputs/predyear_brtHavg_dynsst_mass.RData")
pred_propbiomass_BRTdynsst <- ddply(predyear_brtHavg_dynsst_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

############ TOPOGRAPHIC PREDICTORS ##############
load("./Outputs/predyear_brtHavg_topo_mass.RData")

pred_propbiomass_BRTtopo <- ddply(predyear_brtHavg_topo_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

load("./Outputs/predyear_crossENSavg_topo_mass.RData")

pred_propbiomass_ENStopo <- ddply(predyear_crossENSavg_topo_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})

load("./Outputs/predyear_crossENSavg_topolog_mass.RData")

pred_propbiomass_ENStopolog <- ddply(predyear_crossENSavg_topolog_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})
```

Combine predictions of proportions into a single object
```{r}
pred_propbiomass <- cbind(pred_propbiomass_GAMdynlogdep, 
                          pred_propbiomass_BRTdynlogdep[c("epac_prop", "tspin_prop")], 
                          pred_propbiomass_ENSdynlogdep[c("epac_prop", "tspin_prop")],
                          pred_propbiomass_BRTdynsst[c("epac_prop", "tspin_prop")], 
                          pred_propbiomass_BRTtopo[c("epac_prop", "tspin_prop")],
                          pred_propbiomass_ENStopo[c("epac_prop", "tspin_prop")],
                          pred_propbiomass_ENStopolog[c("epac_prop", "tspin_prop")])
names(pred_propbiomass) <- c("pred_months", "epac_GAMdynlogdep", "tspin_GAMdynlogdep",
                             "epac_BRTdynlogdep", "tspin_BRTdynlogdep", 
                             "epac_ENSdynlogdep", "tspin_ENSdynlogdep", 
                             "epac_BRTdynsst", "tspin_BRTdynsst", 
                             "epac_BRTtopo", "tspin_BRTtopo", 
                             "epac_ENStopo", "tspin_ENStopo",
                             "epac_ENStopolog", "tspin_ENStopolog")

pred_propbiomass <- pred_propbiomass %>% 
  mutate(ras = pmap(list(eg = epac_GAMdynlogdep, tg = tspin_GAMdynlogdep, 
                         eb = epac_BRTdynlogdep, tb = tspin_BRTdynlogdep, 
                         ee = epac_ENSdynlogdep, te = tspin_ENSdynlogdep,
                         ebs = epac_BRTdynsst, tbs = tspin_BRTdynsst, 
                         ebt = epac_BRTtopo, tbt = tspin_BRTtopo,
                         eet = epac_ENStopo, tet = tspin_ENStopo,
                         eetl = epac_ENStopolog, tetl = tspin_ENStopolog), 
                    function(eg, tg, eb, tb, ee, te, ebs, tbs, ebt, tbt, eet, tet, eetl, tetl){
                           r <- stack(list(eg, tg, eb, tb, ee, te, ebs, tbs, ebt, tbt, eet, tet, eetl, tetl))
                           names(r) <- c("epac_GAMdynlogdep", "tspin_GAMdynlogdep", 
                                         "epac_BRTdynlogdep", "tspin_BRTdynlogdep", 
                                         "epac_ENSdynlogdep", "tspin_ENSdynlogdep", 
                                         "epac_BRTdynsst", "tspin_BRTdynsst", 
                                         "epac_BRTtopo", "tspin_BRTtopo", 
                                         "epac_ENStopo", "tspin_ENStopo",
                                         "epac_ENStopolog", "tspin_ENStopolog")
                           return(r)
                         })) %>% 
  mutate(year.month = paste(str_split_i(pred_months, "[.]", 1), 
                            str_split_i(pred_months, "[.]", 2), sep = ".")) %>% 
  select(year.month, ras)

save(pred_propbiomass, file = "./Outputs/pred_propbiomass.RData")
```

## Map biomass proportion

Predicted proportion of species biomass
```{r}
############################################################
## with ENS Biomass dynlogdep ##############################
g_tspinprop <- pred_propbiomass_ENSdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_dynlogdep_ENS_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop

g_epacprop <- pred_propbiomass_ENSdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_dynlogdep_ENS_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop

############################################################
## with BRT Biomass dynlogdep ##############################
g_tspinprop <- pred_propbiomass_BRTdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_dynlogdep_BRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop

g_epacprop <- pred_propbiomass_BRTdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_dynlogdep_BRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop 


############################################################
## with GAM Biomass dynlogdep ##############################
g_tspinprop <- pred_propbiomass_GAMdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_dynlogdep_GAM_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop 

g_epacprop <- pred_propbiomass_GAMdynlogdep %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_dynlogdep_GAM_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop


############################################################
## with BRT Biomass dynsst #################################
g_tspinprop <- pred_propbiomass_BRTdynsst %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_dynsst_BRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop

g_epacprop <- pred_propbiomass_BRTdynsst %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_dynsst_BRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop 

############################################################
## with BRT Biomass topo ###################################
g_tspinprop <- pred_propbiomass_BRTtopo %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_topo_BRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop

g_epacprop <- pred_propbiomass_BRTtopo %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_topo_BRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop 

############################################################
## with ENS Biomass topolog ###################################
g_tspinprop <- pred_propbiomass_ENStopolog %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = tspin_prop) %>% 
  Fun_propmaps()

g_tspinprop <- g_tspinprop + ggtitle("TSPIN biomass proportion")

ggsave(g_tspinprop, file = paste0("./Outputs/BiomassProp_topolog_ENS_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspinprop

g_epacprop <- pred_propbiomass_ENStopolog %>%
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018) %>% 
  mutate(meanpred = epac_prop) %>% 
  Fun_propmaps()

g_epacprop <- g_epacprop + ggtitle("EPAC biomass proportion")

ggsave(g_epacprop, file = paste0("./Outputs/BiomassProp_topolog_ENS_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epacprop 
```

## Calculate NASC x Biomass prop

```{r}
################################################################
########## ENS NASC x ENS Biomass prop Dynlogdep ###############

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSdynlogdep <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_ENSdynlogdep, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSdynlogdep, file = "./Outputs/pred_nascXmass_ENSdynlogdep.RData")

pred_nascXmass_ENSdynlogdep_selec <- pred_nascXmass_ENSdynlogdep %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSdynlogdep_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC x pred biomass proportion")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_dynlogdep_ENS_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSdynlogdep_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC x pred biomass proportion")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_dynlogdep_ENS_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

```{r}
################################################################
########## BRT NASC x BRT Biomass prop Dynlogdep ###############

load("../3-Krill_SDM/Outputs/predyear_brtHavg_dynlogdep.RData")

pred_nascXmass_BRTdynlogdep <- join(predyear_brtHavg_dynlogdep, pred_propbiomass_BRTdynlogdep, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_BRTdynlogdep, file = "./Outputs/pred_nascXmass_BRTdynlogdep.RData")

pred_nascXmass_BRTdynlogdep_selec <- pred_nascXmass_BRTdynlogdep %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_BRTdynlogdep_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC x pred biomass proportion")
g_epac
ggsave(g_epac, file = paste0("./Outputs/nascXmass_dynlogdep_BRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")

g_tspin <- pred_nascXmass_BRTdynlogdep_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC x pred biomass proportion")
g_tspin
ggsave(g_tspin, file = paste0("./Outputs/nascXmass_dynlogdep_BRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
```


```{r}
################################################################
########## ENS NASC x BRT Biomass prop Dynlogdep ###############

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSBRTdynlogdep <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_BRTdynlogdep, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSBRTdynlogdep, file = "./Outputs/pred_nascXmass_ENSBRTdynlogdep.RData")

pred_nascXmass_ENSBRTdynlogdep_selec <- pred_nascXmass_ENSBRTdynlogdep %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSBRTdynlogdep_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC x pred biomass proportion")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_dynlogdep_ENSBRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSBRTdynlogdep_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC x pred biomass proportion")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_dynlogdep_ENSBRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

```{r}
################################################################
########## ENS NASC x BRT Biomass prop Dynsst ###############

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSBRTdynsst <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_BRTdynsst, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSBRTdynsst, file = "./Outputs/pred_nascXmass_ENSBRTdynsst.RData")

pred_nascXmass_ENSBRTdynsst_selec <- pred_nascXmass_ENSBRTdynsst %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSBRTdynsst_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC ENS dynlogdep x pred biomass proportion BRT dynsst")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_dynsst_ENSBRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSBRTdynsst_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC ENS dynlogdep x pred biomass proportion BRT dynsst")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_dynsst_ENSBRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

```{r}
################################################################
########## ENS NASC x GAM Biomass prop Dynlogdep ###############

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSGAMdynlogdep <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_GAMdynlogdep, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSGAMdynlogdep, file = "./Outputs/pred_nascXmass_ENSxGAMdynlogdep.RData")

pred_nascXmass_ENSGAMdynlogdep_selec <- pred_nascXmass_ENSGAMdynlogdep %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSGAMdynlogdep_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC x pred biomass proportion")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_dynlogdep_ENSGAM_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSGAMdynlogdep_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC x pred biomass proportion")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_dynlogdep_ENSGAM_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

```{r}
################################################################
########## ENS NASC x BRT Biomass prop Topo ####################

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSBRTtopo <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_BRTtopo, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSBRTtopo, file = "./Outputs/pred_nascXmass_ENSBRTtopo.RData")

pred_nascXmass_ENSBRTtopo_selec <- pred_nascXmass_ENSBRTtopo %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSBRTtopo_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC dynlogdep ENS x pred biomass proportion topo BRT")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_topo_ENSBRT_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSBRTtopo_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC dynlogdep ENS x pred biomass proportion topo BRT")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_topo_ENSBRT_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```


```{r}
################################################################
########## ENS NASC x ENS Biomass prop Topo ####################

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSENStopo <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_ENStopo, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSENStopo, file = "./Outputs/pred_nascXmass_ENSENStopo.RData")

pred_nascXmass_ENSENStopo_selec <- pred_nascXmass_ENSENStopo %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSENStopo_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC dynlogdep ENS x pred biomass proportion topo ENS")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_topo_ENSENS_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSENStopo_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC dynlogdep ENS x pred biomass proportion topo ENS")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_topo_ENSENS_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

```{r}
################################################################
########## ENS NASC x ENS Biomass prop Topolog #################

load("../3-Krill_SDM/Outputs/predyear_crossENSavg_dynlogdep.RData")

pred_nascXmass_ENSENStopolog <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_ENStopolog, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSENStopolog, file = "./Outputs/pred_nascXmass_ENSENStopolog.RData")

pred_nascXmass_ENSENStopolog_selec <- pred_nascXmass_ENSENStopolog %>% 
  filter(substr(pred_months, 7, 7) %in% c("2","5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018)

# map predictions for three months of interest
g_epac <- pred_nascXmass_ENSENStopolog_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac <- g_epac + ggtitle("EPAC: pred NASC dynlogdep ENS x pred biomass proportion topolog ENS")

ggsave(g_epac, file = paste0("./Outputs/nascXmass_topolog_ENSENS_EPAC.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_epac

g_tspin <- pred_nascXmass_ENSENStopolog_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin <- g_tspin + ggtitle("TSPIN: pred NASC dynlogdep ENS x pred biomass proportion topolog ENS")

ggsave(g_tspin, file = paste0("./Outputs/nascXmass_topolog_ENSENS_TSPIN.png"), width = 180, height = 180, dpi = 500, units = "mm")
g_tspin
```

## Evaluate proportion predictions

Create function to generate plots of observed vs predicted proportion
```{r}
Fun_evalprop <- function(d, x, y, method, sp){
  d$x2 <- unlist(d[x]/d$mass_tot)
  ggpubr::ggscatter(d, x = "x2", y= y, add = "reg.line") +
    ggpubr::stat_cor(label.y = 1) +
    ggpubr::stat_regline_equation(label.y = 0.8) +
    mon_theme +
    ggtitle(method) +
    xlab(paste0("Observed proportion of ", sp)) +
    ylab(paste0("Predicted proportion of ", sp)) +
    ylim(0, 1) +
    xlim(0, 1)
}
```

Explore NH Line distribution
```{r}
mass_nhl <- biomass_nhl_df %>% 
  select(Station, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot) %>% 
  pivot_longer(cols = c(mass_EPAC, mass_TSPIN), names_to = "species", values_to = "mass", names_prefix = "mass_")

g1 <- ggplot(mass_nhl, aes(Station, log(mass + 1), fill = species)) + 
  geom_boxplot()

prop_nhl <- biomass_nhl_df %>% 
  mutate(prop_EPAC = mass_EPAC / mass_tot,
         prop_TSPIN = mass_TSPIN / mass_tot) %>% 
  select(Station, Sample.Date, prop_EPAC, prop_TSPIN, mass_tot) %>% 
  pivot_longer(cols = c(prop_EPAC, prop_TSPIN), names_to = "species", values_to = "prop", names_prefix = "prop_")

g2 <- ggplot(prop_nhl, aes(Station, prop, fill = species)) + 
  geom_boxplot()

g <- g1 + g2 &
  scale_fill_manual(name = "Species", values = krillsp_pal) &
  mon_theme

ggsave(g, file = "./Outputs/NHLine_biomass.png", width = 180, height = 90, dpi = 300, units = "mm")
```

Create dataset of evaluation with NH line and NCC stations
```{r}
biomass_nhl_df$program <- "NHL"
biomass_df$program <- "NCC"

all_biomass_eval_df <- rbind(biomass_nhl_df %>% 
             select(program, date_raster_week, Station, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, utmx, utmy),
                             biomass_df %>% 
             select(program, date_raster_week, Station, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, utmx, utmy)) %>% 
  mutate(year.month = paste(str_split_i(date_raster_week, "[.]", 1), str_split_i(date_raster_week, "[.]", 2), sep = ".")) %>% 
  ddply(., ~year.month, function(d){ # divide biomass data by month
    # select predictions for that month
    pred_propbiomass_selec <- pred_propbiomass %>% filter(year.month == d$year.month[1])
    # extract epac and tspin proportions calculated with gam, brt, ens, dynlogdep or topo
    d[c("epac_GAMdynlogdep", "tspin_GAMdynlogdep", 
                                         "epac_BRTdynlogdep", "tspin_BRTdynlogdep", 
                                         "epac_ENSdynlogdep", "tspin_ENSdynlogdep", 
                                         "epac_BRTdynsst", "tspin_BRTdynsst", 
                                         "epac_BRTtopo", "tspin_BRTtopo", 
                                         "epac_ENStopo", "tspin_ENStopo",
                                         "epac_ENStopolog", "tspin_ENStopolog")] <- raster::extract(x = pred_propbiomass_selec$ras[[1]], y = d[c("utmx", "utmy")])
    return(d)
  }) %>% 
  filter(mass_tot != 0) # remove stations at which there was absolutely no krill (15 stations in NH line)

save(all_biomass_eval_df, file = "./Outputs/all_biomass_eval_df.RData")
```

### Evaluate vs NH line

```{r}
load("./Outputs/pred_propbiomass.RData")

a1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_TSPIN", y = "tspin_ENSdynlogdep", method = "ENS - dynlogdep", sp = "Tspin")
b1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_EPAC", y = "epac_ENSdynlogdep", method = "", sp = "Epac")

a2 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_TSPIN", y = "tspin_BRTdynlogdep", method = "BRT - dynlogdep", sp = "Tspin")
b2 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_EPAC", y = "epac_BRTdynlogdep", method = "", sp = "Epac")

a3 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_TSPIN", y = "tspin_GAMdynlogdep", method = "GAM - dynlogdep", sp = "Tspin")
b3 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_EPAC", y = "epac_GAMdynlogdep", method = "", sp = "Epac")

g <- (a1 + b1) / (a2 + b2) / (a3 + b3)
ggsave(g, file = paste0("./Outputs/eval_NHL_prop_biomass.png"), width = 180, height = 200, dpi = 500, units = "mm")
g

nrow(subset(all_biomass_eval_df, program == "NHL"))

# map example
bio_selec <- all_biomass_eval_df %>% filter(program == "NHL", year.month == "X2020.9") %>% 
  st_as_sf(coords = c("utmx", "utmy"), crs = 32610)
pred_propbiomass_selec <- pred_propbiomass %>% filter(year.month == "X2020.9") %>% pull(ras)
pred_selec <- as.data.frame(pred_propbiomass_selec[[1]])
pred_selec[c("utmx", "utmy")] <- coordinates(pred_propbiomass_selec[[1]])

g <- ggplot() +
  geom_tile(data = pred_selec, aes(utmx, utmy, fill = epac_ENSdynlogdep)) +
  scale_fill_viridis() +
  geom_sf(data = bio_selec, aes(geometry = geometry)) +
  ggtitle("X2020.9 EPAC prop pred vs NH line") +
  theme(axis.text = element_blank())
ggsave(g, file = paste0("./Outputs/eval_NHL_prop_example.png"), width = 180, height = 180, dpi = 300, units = "mm")
g
```

Dynsst models
```{r}
load("./Outputs/pred_propbiomass.RData")

a1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_TSPIN", y = "tspin_BRTdynsst", method = "BRT - dynsst", sp = "Tspin")
b1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NHL"), 
                   x = "mass_EPAC", y = "epac_BRTdynsst", method = "", sp = "Epac")

g <- (a1 + b1) 
ggsave(g, file = paste0("./Outputs/eval_NHL_prop_biomass_dynsst.png"), width = 180, height = 67, dpi = 500, units = "mm")
g
```

### Evaluate vs NCC cruise stations

```{r}
load("./Outputs/pred_propbiomass.RData")

a1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_TSPIN", y = "tspin_ENSdynlogdep", method = "ENS - dynlogdep", sp = "Tspin")
b1 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_EPAC", y = "epac_ENSdynlogdep", method = "", sp = "Epac")

a2 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_TSPIN", y = "tspin_BRTdynlogdep", method = "BRT - dynlogdep", sp = "Tspin")
b2 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_EPAC", y = "epac_BRTdynlogdep", method = "", sp = "Epac")

a3 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_TSPIN", y = "tspin_GAMdynlogdep", method = "GAM - dynlogdep", sp = "Tspin")
b3 <- Fun_evalprop(subset(all_biomass_eval_df, program == "NCC"), 
                   x = "mass_EPAC", y = "epac_GAMdynlogdep", method = "", sp = "Epac")

g <- (a1 + b1) / (a2 + b2) / (a3 + b3)
ggsave(g, file = paste0("./Outputs/eval_NHL_prop_biomass.png"), width = 180, height = 200, dpi = 500, units = "mm")
g

nrow(subset(all_biomass_eval_df, program == "NCC"))

# map example
bio_selec <- all_biomass_eval_df %>% filter(program == "NCC", year.month == "X2020.9") %>% 
  st_as_sf(coords = c("utmx", "utmy"), crs = 32610)
pred_propbiomass_selec <- pred_propbiomass %>% filter(year.month == "X2020.9") %>% pull(ras)
pred_selec <- as.data.frame(pred_propbiomass_selec[[1]])
pred_selec[c("utmx", "utmy")] <- coordinates(pred_propbiomass_selec[[1]])

g <- ggplot() +
  geom_tile(data = pred_selec, aes(utmx, utmy, fill = epac_ENSdynlogdep)) +
  scale_fill_viridis() +
  geom_sf(data = bio_selec, aes(geometry = geometry)) +
  ggtitle("X2020.9 EPAC prop pred vs NH line") +
  theme(axis.text = element_blank())
ggsave(g, file = paste0("./Outputs/eval_NCC_prop_example.png"), width = 180, height = 180, dpi = 300, units = "mm")
g
```


### Evaluate vs NCC and NH Line data combined

Dynlogdep models
```{r}
load("./Outputs/pred_propbiomass.RData")

a1 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_TSPIN", y = "tspin_ENSdynlogdep", method = "ENS - dynlogdep", sp = "Tspin")
b1 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_EPAC", y = "epac_ENSdynlogdep", method = "", sp = "Epac")

a2 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_TSPIN", y = "tspin_BRTdynlogdep", method = "BRT - dynlogdep", sp = "Tspin")
b2 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_EPAC", y = "epac_BRTdynlogdep", method = "", sp = "Epac")

a3 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_TSPIN", y = "tspin_GAMdynlogdep", method = "GAM - dynlogdep", sp = "Tspin")
b3 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_EPAC", y = "epac_GAMdynlogdep", method = "", sp = "Epac")

g <- (a1 + b1) / (a2 + b2) / (a3 + b3)
ggsave(g, file = paste0("./Outputs/eval_all_prop_biomass.png"), width = 180, height = 200, dpi = 500, units = "mm")
g
```

Dynsst models
```{r}
load("./Outputs/pred_propbiomass.RData")

a1 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_TSPIN", y = "tspin_BRTdynsst", method = "BRT - dynsst", sp = "Tspin")
b1 <- Fun_evalprop(all_biomass_eval_df, 
                   x = "mass_EPAC", y = "epac_BRTdynsst", method = "", sp = "Epac")

g <- (a1 + b1) 
ggsave(g, file = paste0("./Outputs/eval_all_prop_biomass_dynsst.png"), width = 180, height = 67, dpi = 500, units = "mm")
g
```

Topographic model
```{r}
# BRT topo
a1 <- Fun_evalprop(all_biomass_eval_df, x = "mass_TSPIN", y = "tspin_BRTtopo", method = "BRT - topo", sp = "Tspin")
b1 <- Fun_evalprop(all_biomass_eval_df, x = "mass_EPAC", y = "epac_BRTtopo", method = "", sp = "Epac")

# ENS topo
a2 <- Fun_evalprop(all_biomass_eval_df, x = "mass_TSPIN", y = "tspin_ENStopo", method = "ENS - topo", sp = "Tspin")
b2 <- Fun_evalprop(all_biomass_eval_df, x = "mass_EPAC", y = "epac_ENStopo", method = "", sp = "Epac")

# ENS topo log
a3 <- Fun_evalprop(all_biomass_eval_df, x = "mass_TSPIN", y = "tspin_ENStopolog", method = "ENS - topolog", sp = "Tspin")
b3 <- Fun_evalprop(all_biomass_eval_df, x = "mass_EPAC", y = "epac_ENStopolog", method = "", sp = "Epac")

g <-  (a1 + b1) / (a2 + b2) / (a3 + b3)
ggsave(g, file = paste0("./Outputs/eval_all_prop_biomass_topo.png"), width = 180, height = 200, dpi = 500, units = "mm")
```
Note: a lot of high observed Epac proportion (bottom right of right panel) are predicted to have low epac proportion in this model

# Trends in final relative abundance by species

## Temporal trends across the whole study region

ENS nasc x ENS mass dynlogdep
```{r}
load("./Outputs/pred_nascXmass_ENSdynlogdep.RData")

krill_final_ENSxENS_trend <- pred_nascXmass_ENSdynlogdep %>%
  mutate(sum_abu_epac = map_dbl(pred_epac, function(r){
    sum(values(r), na.rm = T)
  })) %>%
  mutate(sum_abu_tspin = map_dbl(pred_tspin, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(pred_months, sum_abu_epac, sum_abu_tspin) %>% 
  pivot_longer(cols = c(sum_abu_epac, sum_abu_tspin), 
               names_to = "group", values_to = "sum_abu",
               names_prefix = "sum_abu_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_final_ENSxENS_trend, file = "./Outputs/krill_final_ENSxENS_trend.RData")

# plots trends by year
g <- ggplot(krill_final_ENSxENS_trend, aes(month, sum_abu, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
ggsave(g, file = "./Outputs/pred_nascxmass_ENSxENSdynlogdep_trend.png", width = 180, height = 70, dpi = 500, units = "mm")
g
```

ENS nasc x ENS mass topolog
```{r}
load("./Outputs/pred_nascXmass_ENSENStopolog.RData")

krill_final_ENSxENStopolog_trend <- pred_nascXmass_ENSENStopolog %>%
  mutate(sum_abu_epac = map_dbl(pred_epac, function(r){
    sum(values(r), na.rm = T)
  })) %>%
  mutate(sum_abu_tspin = map_dbl(pred_tspin, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(pred_months, sum_abu_epac, sum_abu_tspin) %>% 
  pivot_longer(cols = c(sum_abu_epac, sum_abu_tspin), 
               names_to = "group", values_to = "sum_abu",
               names_prefix = "sum_abu_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_final_ENSxENStopolog_trend, file = "./Outputs/krill_final_ENSxENStopolog_trend.RData")

# plots trends by year
g <- ggplot(krill_final_ENSxENStopolog_trend, aes(month, sum_abu, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
ggsave(g, file = "./Outputs/pred_nascxmass_ENSxENStopolog_trend.png", width = 180, height = 70, dpi = 500, units = "mm")
g
```

Mean proportion
```{r}
load("./Outputs/krill_final_ENSxENStopolog_trend.RData")

prop_sp <- krill_final_ENSxENStopolog_trend %>% 
  group_by(group) %>% 
  summarize(mean_abu = mean(sum_abu))

prop_sp[prop_sp$group == "tspin", "mean_abu"] / prop_sp[prop_sp$group == "epac", "mean_abu"]
```

ENS nasc x BRT biomass
```{r}
load("./Outputs/pred_nascXmass_ENSBRTdynlogdep.RData")

krill_final_ENSxBRT_trend <- pred_nascXmass_ENSBRTdynlogdep %>%
  mutate(sum_abu_epac = map_dbl(pred_epac, function(r){
    sum(values(r), na.rm = T)
  })) %>%
  mutate(sum_abu_tspin = map_dbl(pred_tspin, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(pred_months, sum_abu_epac, sum_abu_tspin) %>% 
  pivot_longer(cols = c(sum_abu_epac, sum_abu_tspin), 
               names_to = "group", values_to = "sum_abu",
               names_prefix = "sum_abu_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_final_ENSxBRT_trend, file = "./Outputs/krill_final_ENSxBRT_trend.RData")

# plots trends by year
g <- ggplot(krill_final_ENSxBRT_trend, aes(month, sum_abu, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
ggsave(g, file = "./Outputs/pred_nascxmass_ENSxBRTdynlogdep_trend.png", width = 180, height = 70, dpi = 500, units = "mm")
g
```

Mean proportion
```{r}
load("./Outputs/krill_final_ENSxBRT_trend.RData")

prop_sp <- krill_final_ENSxBRT_trend %>% 
  group_by(group) %>% 
  summarize(mean_abu = mean(sum_abu))

prop_sp[prop_sp$group == "tspin", "mean_abu"] / prop_sp[prop_sp$group == "epac", "mean_abu"]
```

ENS nasc x BRT biomass topo
```{r}
load("./Outputs/pred_nascXmass_ENSBRTtopo.RData")

krill_final_ENSxBRTtopo_trend <- pred_nascXmass_ENSBRTtopo %>%
  mutate(sum_abu_epac = map_dbl(pred_epac, function(r){
    sum(values(r), na.rm = T)
  })) %>%
  mutate(sum_abu_tspin = map_dbl(pred_tspin, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(pred_months, sum_abu_epac, sum_abu_tspin) %>% 
  pivot_longer(cols = c(sum_abu_epac, sum_abu_tspin), 
               names_to = "group", values_to = "sum_abu",
               names_prefix = "sum_abu_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_final_ENSxBRTtopo_trend, file = "./Outputs/krill_final_ENSxBRTtopo_trend.RData")

# plots trends by year
g <- ggplot(krill_final_ENSxBRTtopo_trend, aes(month, sum_abu, color = year, group = year)) +
  geom_point() +
  geom_path() +
  scale_y_sqrt("Summed predicted biomass across the study area") +
  mon_theme +
  facet_grid(~group) +
  theme(axis.text.x = element_text(angle = 45))
ggsave(g, file = "./Outputs/pred_nascxmass_ENSxBRTtopo_trend.png", width = 180, height = 70, dpi = 500, units = "mm")
g
```

Mean proportion
```{r}
load("./Outputs/krill_final_ENSxBRTtopo_trend.RData")

prop_sp <- krill_final_ENSxBRTtopo_trend %>% 
  group_by(group) %>% 
  summarize(mean_abu = mean(sum_abu))

prop_sp[prop_sp$group == "tspin", "mean_abu"] / prop_sp[prop_sp$group == "epac", "mean_abu"]
```



## Temporal trends along NH Line

Create a 5 km high raster that just covers the NH line
```{r}
# depth raster
r <- env_pred_full$env_ras[[1]]$data_pred[[1]]$DEPTH

# mean latitude of NH Line
nhl_lat <- mean(biomass_nhl_df$utmy)
# creating a 5 km band around that latitude
extent_nhl_band <-  extent(extent(r)@xmin, extent(r)@xmax, nhl_lat - 2500, nhl_lat + 2500)

# extract depth along the same line
depth_nhl <- crop(r, extent_nhl_band)
```

### with ENS NASC x ENS biomass dynlogdep

```{r}
load("./Outputs/pred_nascXmass_ENSdynlogdep.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSxENS_trend <- pred_nascXmass_ENSdynlogdep %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSxENS_trend, file = "./Outputs/krill_nhl_pred_ENSxENS_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSxENS_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with longitude (sum by latitude)
krill_nhl_pred_ENSxENS_trend %>% 
  group_by(group, year, utmx) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, utmx) %>% 
  ggplot(., aes(utmx, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

## cross-shelf trend with depth (sum by latitude)
krill_nhl_pred_ENSxENS_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(-DEPTH, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_x_log10() +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
```

### with ENS NASC x BRT biomass dynlogdep
```{r}
load("./Outputs/pred_nascXmass_ENSBRTdynlogdep.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSxBRT_trend <- pred_nascXmass_ENSBRTdynlogdep %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSxBRT_trend, file = "./Outputs/krill_nhl_pred_ENSxBRT_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSxBRT_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with longitude (sum by latitude)
krill_nhl_pred_ENSxBRT_trend %>% 
  group_by(group, year, utmx) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, utmx) %>% 
  ggplot(., aes(utmx, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

## cross-shelf trend with depth (sum by latitude)
g1 <- krill_nhl_pred_ENSxBRT_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(mean_pred = mean(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(log(-DEPTH + 1), mean_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Mean predicted biomass along NH Line") +
    scale_x_continuous(limits = log(c(10, max(-krill_nhl_pred_ENSxBRT_trend$DEPTH))),
                       breaks = log(c(10, 50, 100, 200, 500, 1000, 2000)),
                       labels = c(10, 50, 100, 200, 500, 1000, 2000)) +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

g2 <- biomass_nhl_df %>% 
  select(Station, Year, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, DEPTH) %>% 
  pivot_longer(cols = c(mass_EPAC, mass_TSPIN), names_to = "group", values_to = "mass", names_prefix = "mass_") %>% 
  group_by(group, Year, DEPTH) %>% 
  mutate(mean_obs = mean(mass, na.rm = T)) %>% 
  ggplot(., aes(log(-DEPTH + 1), mean_obs, color = group)) +
    geom_point(show.legend = F) +    
    scale_y_sqrt("Observed biomass along NH Line") +
    scale_x_continuous(limits = log(c(10, max(-krill_nhl_pred_ENSxBRT_trend$DEPTH))),
                       breaks = log(c(10, 50, 100, 200, 500, 1000, 2000)),
                       labels = c(10, 50, 100, 200, 500, 1000, 2000)) +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
g <- g1 / g2
ggsave(g, file = "./Outputs/krill_nhl_pred_ENSxBRT_trend_depth.png", width = 140, height = 120, units = "mm", dpi = 300)
g
```
Note to myself: ENSxBRT models get the depth gradient in Epac / Tspin better with more tspin than epac within 200 m of depth. But the brt therefore tends to inflate a bit the overall abundance of tspin and the summed abndance through time therefore do not look at much of a Tspin / Epac difference than we had with the ENS x ENS approach (whereby epac was >> to tspin over all months)


### with ENS NASC x BRT biomass topo
```{r}
load("./Outputs/pred_nascXmass_ENSBRTtopo.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSxBRTtopo_trend <- pred_nascXmass_ENSBRTtopo %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSxBRTtopo_trend, file = "./Outputs/krill_nhl_pred_ENSxBRTtopo_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSxBRTtopo_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with depth (sum by latitude)
g1 <- krill_nhl_pred_ENSxBRTtopo_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(mean_pred = mean(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(log(-DEPTH + 1), mean_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Mean predicted biomass along NH Line") +
    scale_x_continuous(limits = log(c(10, max(-krill_nhl_pred_ENSxBRTtopo_trend$DEPTH))),
                       breaks = log(c(10, 50, 100, 200, 500, 1000, 2000)),
                       labels = c(10, 50, 100, 200, 500, 1000, 2000)) +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

g2 <- biomass_nhl_df %>% 
  select(Station, Year, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, DEPTH) %>% 
  pivot_longer(cols = c(mass_EPAC, mass_TSPIN), names_to = "group", values_to = "mass", names_prefix = "mass_") %>% 
  group_by(group, Year, DEPTH) %>% 
  mutate(mean_obs = mean(mass, na.rm = T)) %>% 
  ggplot(., aes(log(-DEPTH + 1), mean_obs, color = group)) +
    geom_point(show.legend = F) +    
    scale_y_sqrt("Observed biomass along NH Line") +
    scale_x_continuous(limits = log(c(10, max(-krill_nhl_pred_ENSxBRTtopo_trend$DEPTH))),
                       breaks = log(c(10, 50, 100, 200, 500, 1000, 2000)),
                       labels = c(10, 50, 100, 200, 500, 1000, 2000)) +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
g <- g1 / g2
ggsave(g, file = "./Outputs/krill_nhl_pred_ENSxBRTtopo_trend_depth.png", width = 140, height = 120, units = "mm", dpi = 300)
g
```


### with ENS NASC dynlogdep x ENS biomass topo

```{r}
load("./Outputs/pred_nascXmass_ENSENStopo.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSENStopo_trend <- pred_nascXmass_ENSENStopo %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSENStopo_trend, file = "./Outputs/krill_nhl_pred_ENSENStopo_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSENStopo_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with longitude (sum by latitude)
krill_nhl_pred_ENSENStopo_trend %>% 
  group_by(group, year, utmx) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, utmx) %>% 
  ggplot(., aes(utmx, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

## cross-shelf trend with depth (sum by latitude)
krill_nhl_pred_ENSENStopo_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(-DEPTH, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_x_log10() +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
```

### with ENS NASC dynlogdep x ENS biomass topolog

```{r}
load("./Outputs/pred_nascXmass_ENSENStopolog.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSENStopolog_trend <- pred_nascXmass_ENSENStopolog %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSENStopolog_trend, file = "./Outputs/krill_nhl_pred_ENSENStopolog_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSENStopolog_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with longitude (sum by latitude)
krill_nhl_pred_ENSENStopolog_trend %>% 
  group_by(group, year, utmx) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, utmx) %>% 
  ggplot(., aes(utmx, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme

## cross-shelf trend with depth (sum by latitude)
krill_nhl_pred_ENSENStopolog_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(-DEPTH, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_x_log10() +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
```


Compare longitudinal pattern for ens topolog and brt topo to see which one captured the offshore decline of tspin best
```{r}
rbind(krill_nhl_pred_ENSENStopolog_trend %>% 
        mutate(method = c("ENSdynlogdep X ENStopolog")),
      krill_nhl_pred_ENSxBRTtopo_trend %>% 
        mutate(method = c("ENSdynlogdep X BRTtopolog"))
) %>% 
  group_by(method, group, year, DEPTH) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(-DEPTH, sum_pred, color = group, group = interaction(method, group, year))) +
    geom_point(aes(shape = method)) +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    #scale_x_log10() +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    facet_wrap(~method, ncol = 2) +
    mon_theme
    
```

### with ENS NASC dynlogdep x ENS biomass topolog

```{r}
load("./Outputs/pred_nascXmass_ENSBRTdynsst.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSBRTdynsst_trend <- pred_nascXmass_ENSBRTdynsst %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSBRTdynsst_trend, file = "./Outputs/krill_nhl_pred_ENSBRTdynsst_trend.RData")

# plot predicted species abundance along line, in response to depth and to time

## temporal trend (sum by month)
krill_nhl_pred_ENSBRTdynsst_trend %>% 
  group_by(group, period) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  ggplot(., aes(month, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme +
    theme(axis.text.x = element_text(angle = 45))

## cross-shelf trend with depth (sum by latitude)
krill_nhl_pred_ENSBRTdynsst_trend %>% 
  group_by(group, year, DEPTH) %>% 
  mutate(sum_pred = sum(pred, na.rm = T)) %>% 
  arrange(group, year, DEPTH) %>% 
  ggplot(., aes(-DEPTH, sum_pred, color = group, group = interaction(group, year))) +
    geom_point() +
    geom_path() +
    scale_y_sqrt("Summed predicted biomass along NH Line") +
    scale_x_log10() +
    scale_color_manual(name = "Species", values = krillsp_pal) +
    mon_theme
```



