---
title: "Krill NASC SDM"
author: "Solene Derville"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
   html_document:
    highlight: pygments
    keep_md: false
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
---

# FILE PATHS
```{r, echo=F, result="hide", include=F}
lib=lapply(c("plyr","officer","raster","stars", "sf", "viridis", "ggthemes", "patchwork", "doParallel", "foreach", 
             "tidyverse", "lubridate", "mgcv", "gratia", "ape", "gstat", "RColorBrewer", "caret", "gbm", "dismo", 
             "pROC", "gstat", "future", "future.apply", "furrr", "fmsb", "splitTools"), library, character.only=T)

mon_theme <- theme(
        panel.border = element_rect(linewidth=0.5,color="black", fill="transparent"),
        plot.margin=unit(c(2,2,2,2),"mm"),
        panel.background = element_rect(fill = 'white'),
        text=element_text(face="bold", size=7),
        title = element_text(size=rel(1)))

Sys.setenv(TZ = "America/Los_Angeles")
```

Environmental data for predictions
```{r, warning = F}
load("../Data/Environment/env_pred_full.RData")
depth_ras <- env_pred_full$env_ras[[1]]$data_pred[[1]]$DEPTH
```

Environmental data for maps
```{r, warning = F, message = F}
# coastline shapefile
coast_sf <- st_read("../Data/Environment/coast_openstreetmap_WA-OR-CA.shp")
coast_sf_utm <- st_transform(coast_sf, crs = 32610)

# isobaths
iso_sf <- st_read("../Data/Environment/isobaths_50mto1500m.shp")
iso_sf_utm <- st_transform(iso_sf, crs = 32610)

# isobath 1500 m
iso_sf1500 <- st_read("../Data/Environment/iso_sf1500.shp")

# Extent of study area for predictions (OR coastal waters only)
extent_study_area <-  extent(-125.32, -123, 42, 46.3)
polygon_study_area <- as(extent_study_area, 'SpatialPolygons') %>%
  st_as_sf() %>%
  st_set_crs(value = 4326)
polygon_study_area_utm <- polygon_study_area %>% st_transform(crs = 32610)
```

Echosounder data
```{r}
load("../2-Environment/Outputs/krill_agg_df.RData")
# select only daytime
suppressWarnings(krill_day_df <- krill_agg_df %>% 
  filter(phase == "day") %>% 
  mutate(yday = yday(as.POSIXct(as.character(Date_M), format = "%Y%m%d")),
         CruiseID = paste0("SH", substr(Date_M, 3, 6))) %>% 
  # a bunch of cruise ID need to be recoded to match with those in the bongo data
  # ignore warning
  mutate(CruiseID = car::recode(CruiseID, "'SH1805' = 'SH1804'; 'SH1809'='SH1811'; 'SH1905'='SH1904'; 'SH1909'='SH1908'; 'SH2003'='SH2002'; 'SH2209'='SH2210'; 'SH2009'='SH2010'")) %>% 
  # add log depth
  mutate(logDEPTH = log(-DEPTH)))

nrow(krill_day_df)

save(krill_day_df, file = "./Outputs/krill_day_df.RData")
```

Functions
```{r}
# create function to make dataframe predictions from a hierarchical modelling approach: binomial and gamma models
# to specific periods stored in env layers ras
# m_name is the model name = predictors chosen
# here fam should be a vector of the two families bin and lognasc
# method is to indicate if the model is a GAM or BRT
Fun_Hpredras <- function(m_name, fam, m_tb, ras, met){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep <- tb %>% 
      filter(model_name == m_name & family == fam & method == met) %>% 
      pull(model_var)
    # model of presence
    m1 <- tb %>% 
      filter(model_name == m_name & family == fam[1] & method == met) %>% 
      pull(model)
    load(m1[[1]])
    if(met == "BRT"){m1m <- mod_brt$model[[1]]}
    if(met == "GAM"){m1m <- mod_gam}
    # model of abundance
    m2 <- tb %>% 
      filter(model_name == m_name & family == fam[2] & method == met) %>% 
      pull(model)
    load(m2[[1]])
    if(met == "BRT"){m2m <- mod_brt$model[[1]]}
    if(met == "GAM"){m2m <- mod_gam}
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week <- r_week[[var_tokeep[[1]]]]
            # then make predictions on raster!
            if(met == "BRT"){
              pred1 <- raster::predict(r_week, m1m, type = "response",
                                      n.trees = m1m$gbm.call$best.trees)
              pred2 <-  raster::predict(r_week, m2m, type = "response",
                                      n.trees = m2m$gbm.call$best.trees)
              pred2 <- exp(pred2)} # response is logged in brt
            if(met == "GAM"){
              pred1 <- raster::predict(r_week, m1m, type = "response")
              pred2 <- raster::predict(r_week, m2m, type = "response")
            }
            predH <- pred1 * pred2
            return(predH)
          }))
          # average across four weeks with NA.rm
          mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
          return(mean_pred)
      }))
  })
}

## BRT bin x GAM abun
Fun_crossENSpredras <- function(m_name, m_tb, ras){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep_bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "GAM") %>% 
      pull(model_var)
    var_tokeep_abu <- tb %>% 
      filter(model_name == m_name & family == "lognasc" & method == "GAM") %>% 
      pull(model_var)
    # model of presence
    m.bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "BRT") %>% 
      pull(model)
    load(m.bin[[1]])
    m.bin.m <- mod_brt$model[[1]]
    # model of abundance
    m.abu <- tb %>% 
      filter(model_name == m_name & family == "lognasc" & method == "GAM") %>% 
      pull(model)
    load(m.abu[[1]])
    m.abu.m <- mod_gam
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week_bin <- r_week[[var_tokeep_bin[[1]]]]
            r_week_abu <- r_week[[var_tokeep_abu[[1]]]]
            # then make predictions on raster!
            pred1 <- raster::predict(r_week_bin, m.bin.m, type = "response",
                                      n.trees = m.bin.m$gbm.call$best.trees)
            pred2 <- raster::predict(r_week_abu, m.abu.m, type = "response")
            predH <- pred1 * pred2
            return(predH)
        }))
        # average across four weeks with NA.rm
        mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
        return(mean_pred)
      }))
  })
}

# create function to turn tibble of rasters into one dataframe that can generate a ggplot
Fun_meanpredmaps <- function(ras_tb){
  
  # transform into sf format
  d <- ras_tb %>% 
    mutate(pred_df = map(meanpred, function(r){
      # turn raster to sf object
      d <- r %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
      # add coordinates to dataframe of predictions
      d[c("x", "y")] <- st_coordinates(d)
      return(d)
    })) %>% 
    mutate(month = lubridate::month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE, locale="EN-us"),
           year = lubridate::year(as.POSIXct(pred_months, format = "X%Y.%m.%d")),
           period = paste(month, year)) %>%  
    dplyr::select(period, month, year, pred_df) %>%
    # arrange and mutate factor levels so that plots show up in the right order by months 
    arrange(year, month) %>% 
    mutate(period = factor(period, levels = unique(period))) %>% 
    unnest(pred_df)
  
  # create facetted maps by year in columns and months in rows
  g <- ggplot() +
    geom_tile(data = d, aes(x, y, fill = log(fit + 0.1))) +
    scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog NASC",
                         n.breaks = 6) +  
    geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
    geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
    facet_wrap(~period, ncol = n_distinct(d$year), dir = "v") +
    coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
    mon_theme +
    xlab("") +
    ylab("") +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
    
    return(g)
}

# create custom function to run BRTs in a hierarchical approach
Fun_mybrt <- function(d, v, fam){
  # find what columns the names of the variables correspond to
  var_num <- match(v, names(d))
  var_y <- match("nasc", names(d))
  # set starting learning rate (usually between 0.01 and 0.001)
  lr <- 0.01 

  target.trees <- 0 # resets the target trees , make sure set condition to lower than 1000

  while (target.trees < 1000) {
  
    mod.brt <- gbm.step(d,
                     gbm.x = var_num, # Your predictor variables, referring to columns in your data frame
                     gbm.y = var_y, # Your response variable
                     family = fam,
                     n.trees = 50,
                     bag.fraction = 0.75, # 0.75 is typical for larger dataset, but can increase bag fraction if dataset is small 
                     plot.main = F,   
                     learning.rate = lr,
                     tree.complexity = 2,
                     verbose = F)  
  
    if (object.size(mod.brt) > 0) {    # > 0 = good, check if hasn't crashed
      target.trees <- mod.brt$gbm.call$best.trees
    }
    else {
      target.trees <- 0  #check if model size is zero 
    }
    lr <- lr / 2
  }
  out <- tibble(model = list(mod.brt),
                lr = lr)
  return(out)
}

# create facetted average 2018-2022 maps by months for GAM and BRT
Fun_YearAvgMap <- function(dd, method, group_name){
    g <- ggplot() +
      geom_tile(data = dd, aes(x, y, fill = log(fit + 0.1))) + 
      scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog NASC",
                         n.breaks = 6) + 
      geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
      geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
      facet_wrap(~month, ncol = 6, dir = "h") +
      coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
      mon_theme +
      xlab("") +
      ylab("") +
      theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
      ggtitle(paste(method, group_name))
    ggsave(g, file = paste0("./Outputs/", method, "_dynlogdep_YearRound_2018-2022_", group_name,".png"), width = 230, height = 150, dpi = 500, units = "mm")
}
```

# Running SDMs
## Set up models

### Predictor formulas
```{r}
form_chr <- list(
     "topographic" = " nasc ~
                   s(DEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5)",
     "dynamic_logdep" = "nasc ~ 
                   s(logDEPTH, k = 5) + 
                   s(DIS_CANYON, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(sst, k = 5) + 
                   s(ssh, k = 5) + 
                   s(eke, k = 5) +
                   s(sshsd, k = 5) + 
                   s(sstsd, k = 5) + 
                   s(curl, k = 5) + 
                   s(ild_05, k = 5) + 
                   s(bbv_200, k = 5)")
model_type <- c("topo", "dyn")
model_var <- list(c("DEPTH","SLOPE", "DIS_CANYON"),
                  c("logDEPTH","DIS_CANYON", "SLOPE", "sst", "ssh", "sshsd","sstsd","eke", "curl", "ild_05", "bbv_200"))
```

### Folds and model types
```{r}
load("./Outputs/krill_day_df.RData")

Fun_folding <- function(df){
  # set seed to NULL so that we can rerun multiple times until we get a combination of folds without zero presence
  folds <- splitTools::create_folds(df$year_month, k = 9, seed = 1, type = "grouped")
  d_cv <- llply(folds, function(f){
    df$fold <- "test"
    df[f,]$fold <- "train"
    # reset all data points from winter 202003 to be in the train set
    # otherwise there is one run of the model training that will not incorporate any winter data
    df[df$year_month == "202003", ]$fold <- "train"
    # do not keep the fold that has no test data (the one for which the winter cruise was set back to train instead of test)
    if(nrow(subset(df, fold == "test")) > 0){
      return(df)
    }
  })
  # null is returned when if() test above is not met
  d_cv <- discard(d_cv, is.null)
  return(d_cv)
}

krill_day_df <- krill_day_df %>% 
      mutate(year_month = substr(Date_M, 1, 6)) %>% 
      mutate(year_month = ifelse(year_month == 201904, 201905, year_month))
unique(krill_day_df$year_month)

krill_cv <- tibble(folds = paste0("Fold", c(1:8)),
                   data = Fun_folding(krill_day_df))

modelsH_krill_tb <- tibble(model_form = form_chr) %>%
    mutate(model_name = names(form_chr),
         model_type = model_type,
         model_var = model_var,
         model_form = map(model_form, as.formula),
         data = rep(list(krill_cv), length(form_chr))) %>% 
  unnest(cols = c(data))
modelsH_krill_tb
```

## Run models in parallel

```{r, eval = F}
modelsH_krill_tb <-  modelsH_krill_tb %>%
  ## BINOMIAL MODELS
  mutate(bin = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, fo = folds), .f = function(a, b, c, mn, fo){ 
      # create data set with binomial response
      a_modified <- a %>% mutate(nasc = ifelse(nasc == 0, 0, 1))
      # in the case of GAMs, the explanatory variables should not have NAs
      a_modified_gam <- a_modified
    
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified, fold == "train"), 
                           v = b, fam = "bernoulli")
      mod_gam <- gam(c, family = binomial(link = "logit"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_bin_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_bin_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return models in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    }))

modelsH_krill_tb <- modelsH_krill_tb %>%
  ## POSITIVE VALUES MODEL
  mutate(lognasc = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, fo = folds), .f = function(a, b, c, mn, fo){ 
      # create data set with positive nasc response
      a_modified_gam <- a %>% 
                    filter(nasc > 0)
      # in the case of BRT the log should be applied prior to running the model
      a_modified_brt <- a %>% 
                    filter(nasc > 0) %>% 
                    mutate(nasc = log(nasc)) # apply a log link
    
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified_brt, fold == "train"), 
                           v = b, fam = "gaussian")
      mod_gam <- gam(c, family = Gamma(link = "log"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_lognasc_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_lognasc_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return path to model objects in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    
    })) 

# pivot tibble and calculate deviance explained by each model
# in training and in test
modelsH_krill_tb <- modelsH_krill_tb %>%
  pivot_longer(cols = c(bin, lognasc), names_to = "family", values_to = "mod") %>% 
  unnest(cols = c(mod))

# correct data depending on family and method
modelsH_krill_tb <- modelsH_krill_tb %>%
  mutate(data = pmap(list(d = data, v = model_var, fa = family, m = method), 
                     function(d, v, fa, m){
      if(fa == "bin" & m == "BRT"){
        dd <- d %>% mutate(nasc = ifelse(nasc == 0, 0, 1))
      }
      if(fa == "lognasc" & m == "BRT"){
        dd <- d %>% filter(nasc > 0) %>% mutate(nasc = log(nasc))
      }
      if(fa == "bin" & m == "GAM"){
        dd <- d %>% mutate(nasc = ifelse(nasc == 0, 0, 1))
        if(!is.null(v)){dd <- dd[complete.cases(dd[v]), ]}
      }
      if(fa == "lognasc" & m == "GAM"){
        dd <- d %>% filter(nasc > 0)
      }
      return(dd)
  }))

save(modelsH_krill_tb, file = "./Outputs/modelsH_krill_tb.RData")
```

# Model performance

## Performance by method by family
```{r, warning = F, message = F}
load("./Outputs/modelsH_krill_tb.RData")

modelsH_krill_perf <- modelsH_krill_tb %>%
  mutate(dev.exp = pmap_dbl(list(mo = model, me = method), function(mo, me){
    load(mo)
    if(me == "BRT"){
      dev <- round((mod_brt$model[[1]]$self.statistics$mean.null - mod_brt$model[[1]]$cv.statistics$deviance.mean)/mod_brt$model[[1]]$self.statistics$mean.null*100, 1)
    }
    if(me == "GAM"){
      dev <- round(summary(mod_gam)$dev.expl*100, 1)
    }
    rm(mo)
    return(dev)
  })) %>%
  # AUC calculation does not depend on method, but can only be done for the bin models
  mutate(AUC.train = pmap_dbl(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    pred.train <- predict(m, subset(d, fold == "train"), type = "response")
    if(fa == "bin"){
        auc <- round(pROC::roc(subset(d, fold == "train")$nasc, pred.train)$auc, 3)
    } else {auc <- NA}
    return(auc)
  })) %>%
  # For the test fold, we calculate either AUC for the bin model, or pear.cor for the lognasc model
  mutate(metrics.test = pmap(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    pred.test <- predict(m, subset(d, fold == "test"), type = "response")
    if(length(pred.test[!is.na(pred.test)])>2 & 
       nrow(subset(d, fold == "test" & nasc != 0)) > 1){ # if too many NAs (e.g. in the chla model), then return NA as metrics
      # and if the test folds contain at least one positive value
      if(fa == "bin"){
        # no tspin in test fold07 so cannot compute metrics when response is all zeros
        auc <- round(pROC::roc(subset(d, fold == "test")$nasc, pred.test)$auc, 3)
        spea.cor <- NA
        pear.cor <- NA
        
      } else {
        auc <- NA
        spea.cor <- round(cor.test(subset(d, fold == "test")$nasc, pred.test, method = "spearman")$estimate, 3)
        pear.cor <- round(cor.test(subset(d, fold == "test")$nasc, pred.test, method = "pearson")$estimate, 3)
      }
    } else { spea.cor <- NA ; auc <- NA ; pear.cor <- NA}
    metrics <- tibble(spea.cor = spea.cor, pear.cor = pear.cor, AUC.test = auc)
    return(metrics)
  })) %>% 
  unnest(cols = c(metrics.test)) %>% 
  dplyr::select(model_name, family, method, folds, dev.exp, AUC.train, AUC.test, pear.cor, spea.cor) %>% 
  arrange(family)

(mean_perf <- modelsH_krill_perf %>% 
  group_by(model_name, family, method) %>% 
  summarize(mean.dev.exp = mean(dev.exp, na.rm = T),
            mean.AUC.train =  mean(AUC.train, na.rm = T),
            mean.AUC.test = mean(AUC.test, na.rm = T),
            mean.pear.cor = mean(pear.cor, na.rm = T),
            mean.spea.cor = mean(spea.cor, na.rm = T)) %>% 
  mutate_if(is.numeric, ~ round(., digits = 3)) %>% 
  select(family, everything()) %>% 
  arrange(family, -mean.dev.exp))
```

## Select model
Selecting the optimal model
```{r, eval = F}
load("./Outputs/modelsH_krill_tb.RData")
     
modelsH_krill_dynlogdep <- modelsH_krill_tb %>% 
  filter(model_name == "dynamic_logdep")

save(modelsH_krill_dynlogdep, file = "./Outputs/modelsH_krill_dynlogdep.RData")
```

# Maps of predictions

Months x years to predict year-round
```{r}
# select combination of months and years of interest
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full environmental data rasters                 
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)

# check that we have the right months represented
unique(env_pred_yearround$pred_months)
```

Generate predictions from the Ensemble model
```{r, eval = F}
load("./Outputs/modelsH_krill_dynlogdep.RData")

predyear_crossENS_dynlogdep <- Fun_crossENSpredras(m_name = "dynamic_logdep",  m_tb = modelsH_krill_dynlogdep, ras = env_pred_yearround)
save(predyear_crossENS_dynlogdep, file = "./Outputs/predyear_crossENS_dynlogdep.RData")

predyear_crossENSavg_dynlogdep <- ddply(predyear_crossENS_dynlogdep, .(pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_crossENSavg_dynlogdep, file = "./Outputs/predyear_crossENSavg_dynlogdep.RData")
```

# Relationships

## Variable influence
```{r}
load("./Outputs/modelsH_krill_dynlogdep.RData")

## GAM
Fun_VarPVGAM <- function(m){
  load(m)
  data.frame(predvar = names(summary(mod_gam)$chi.sq),
             pv = summary(mod_gam)$s.pv)
}

Fun_VarInfluGAM <- function(d){
  data.frame(count0.001 = length(d$pv[d$pv <= 0.001]),
             count0.01 = length(d$pv[d$pv <= 0.01]),
             count0.05 = length(d$pv[d$pv <= 0.05]),
             mean_pvalue = round(mean(d$pv), 3))
}

influ_gam_dynlogdep <- modelsH_krill_dynlogdep %>% 
  filter(method == "GAM") %>% 
  mutate(influence = map(model, function(m) Fun_VarPVGAM(m))) %>% 
  select(method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  ddply(., .(method, family, predvar), function(d) Fun_VarInfluGAM(d))

  
save(influ_gam_dynlogdep, file = "./Outputs/influ_gam_dynlogdep.RData")


### BRT
influ_brt_dynlogdep <- modelsH_krill_dynlogdep %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)

save(influ_brt_dynlogdep, file = "./Outputs/influ_brt_dynlogdep.RData")

```

## Partial Dependence plots

```{r}
load("./Outputs/krill_day_df.RData")

# create a dataframe filled with the average of each variable
suppressWarnings(
mean_pred_df <- krill_day_df %>% 
    dplyr::select(modelsH_krill_dynlogdep$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))
)

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
ENScurves_krill_dynlogdep <- modelsH_krill_dynlogdep %>% 
  filter((method == "BRT" & family == "bin") | (method == "GAM" & family == "lognasc")) %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    if(me == "BRT"){
      m <- mod_brt$model[[1]]
      var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    }
    if(me == "GAM"){
      m <- mod_gam
      var_selec <- data.frame(predvar = names(summary(m)$chi.sq),
                        pv = summary(m)$s.pv) %>% 
        filter(pv <= 0.05) %>% 
        mutate(var = substr(predvar, 3, (nchar(predvar)-1)))
    }
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(krill_day_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        if(me == "BRT"){
          pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
          pred$se.fit <- NA # no error estimate in the BRT
        }
        if(me == "GAM"){
          pred <- dismo::predict(m, mean_pred_x_df, type = "response", se.fit = T)
        }
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

ENScurves_krill_dynlogdep <- ENScurves_krill_dynlogdep %>% 
  select(family, folds, pred) %>% 
  unnest(cols = c(pred))

save(ENScurves_krill_dynlogdep, file = "./Outputs/ENScurves_krill_dynlogdep.RData")
```

