---
title: "Krill TSPIN and EPAC bongo net biomass SDMs"
author: "Solene Derville"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
   html_document:
    highlight: pygments
    keep_md: false
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
---

### FILE PATHS
```{r, echo=F, result="hide", include=F}
lib=lapply(c("plyr","officer","raster","stars", "sf", "viridis", "ggthemes", "patchwork", "doParallel", "foreach", 
             "tidyverse", "lubridate", "mgcv", "gratia", "ape", "gstat", "RColorBrewer", "caret", "gbm", "dismo", 
             "pROC", "gstat", "future", "future.apply", "furrr", "fmsb", "gridExtra"), library, character.only=T)

mon_theme <- theme(
        panel.border = element_rect(linewidth=0.5,color="black", fill="transparent"),
        plot.margin=unit(c(2,2,2,2),"mm"),
        panel.background = element_rect(fill = 'white'),
        text=element_text(face="bold", size=7),
        title = element_text(size=rel(1)))

krillsp_col <- data.frame(group = c("EPAC", "TSPIN"),
                          col = c("#e1c340", "#4cd7d0"))
krillsp_pal <- c("#e1c340",  "#4cd7d0")

Sys.setenv(TZ = "America/Los_Angeles")
```

Environmental data for predictions
```{r, warning = F}
load("../Data/Environment/env_pred_full.RData")
depth_ras <- env_pred_full$env_ras[[1]]$data_pred[[1]]$DEPTH
```

Environmental data for maps
```{r, warning = F, message = F}
# coastline shapefile
coast_sf <- st_read("../Data/Environment/coast_openstreetmap_WA-OR-CA.shp")
coast_sf_utm <- st_transform(coast_sf, crs = 32610)

# isobaths
iso_sf <- st_read("../Data/Environment/isobaths_50mto1500m.shp")
iso_sf_utm <- st_transform(iso_sf, crs = 32610)

# isobath 1500 m
iso_sf1500 <- st_read("../Data/Environment/iso_sf1500.shp")

# Extent of study area for predictions (OR coastal waters only)
extent_study_area <-  extent(-125.32, -123, 42, 46.3)
extent_study_area_wide <-  extent(-129, -122.3, 37.5, 48) # min and max long correspond to min and max of krill_day_df

# Habitat shapefiles
load("../Data/Environment/zones_split.RData")
```

Bongo nets' data
```{r}
load("../2-Environment/Outputs/biomass_df.RData")
nrow(biomass_df)

###############################
####### NCC cruise stations ###
biomass_df <- biomass_df %>% 
  # remove daytime bongos except for 092118NH65 because it was dusk
  filter(Day_Night != "D" | Sample.Code == "092118NH65") %>% 
  # change names of columns and compute density and mass by species
  rename(mass_EPAC_juv = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Juvenile",
         mass_EPAC_adu = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Adult",
         mass_EPAC_adf = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Female..Adult",
         mass_EPAC_adm = "Biomass.Carbon..mg.m3..EUPHAUSIA.PACIFICA_Male..Adult",
         
         mass_TSPIN_juv = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Juvenile",
         mass_TSPIN_adu = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Adult",
         mass_TSPIN_adf = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Female..Adult",
         mass_TSPIN_adm = "Biomass.Carbon..mg.m3..THYSANOESSA.SPINIFERA_Male..Adult",
         
         dens_EPAC_juv = "Density....m3..EUPHAUSIA.PACIFICA_Juvenile",
         dens_EPAC_adu = "Density....m3..EUPHAUSIA.PACIFICA_Adult",
         dens_EPAC_adf = "Density....m3..EUPHAUSIA.PACIFICA_Female..Adult",
         dens_EPAC_adm = "Density....m3..EUPHAUSIA.PACIFICA_Male..Adult",
         
         dens_TSPIN_juv = "Density....m3..THYSANOESSA.SPINIFERA_Juvenile",
         dens_TSPIN_adu = "Density....m3..THYSANOESSA.SPINIFERA_Adult",
         dens_TSPIN_adf = "Density....m3..THYSANOESSA.SPINIFERA_Female..Adult",
         dens_TSPIN_adm = "Density....m3..THYSANOESSA.SPINIFERA_Male..Adult") %>% 
  mutate(mass_EPAC = mass_EPAC_juv + mass_EPAC_adu + mass_EPAC_adf + mass_EPAC_adm,
         mass_TSPIN = mass_TSPIN_juv + mass_TSPIN_adu + mass_TSPIN_adf + mass_TSPIN_adm,
         mass_tot = mass_EPAC + mass_TSPIN,
         dens_tot = dens_EPAC_adu + dens_EPAC_juv + dens_EPAC_adf + dens_EPAC_adm +
                    dens_TSPIN_adu + dens_TSPIN_juv + dens_TSPIN_adf + dens_TSPIN_adm) %>% 
  st_as_sf() %>% 
  st_join(zones_split) %>% 
  # remove geometry because it makes the brt / gam loop crash
  as.data.frame() %>% 
  select(-geometry) %>% 
  # add log depth
  mutate(logDEPTH = log(-DEPTH))
nrow(biomass_df)

save(biomass_df, file = "./Outputs/biomass_df.RData")

###############################
####### NH Line stations ######
load("../2-Environment/Outputs/biomass_nhl_df.RData")

biomass_nhl_df <- biomass_nhl_df %>% 
  rename(mass_EPAC = EUPHAUSIA.PACIFICA, mass_TSPIN = THYSANOESSA.SPINIFERA) %>% 
  mutate(mass_tot = mass_EPAC + mass_TSPIN)

save(biomass_nhl_df, file = "./Outputs/biomass_nhl_df.RData")
```
Predictions from step 3-NASC_models
```{r}
load("../3-NASC_models/Outputs/predyear_crossENSavg_dynlogdep.RData")
```

Functions
```{r}
# create custom function to run BRTs in a hierarchical approach
Fun_mybrt <- function(d, v, fam){
  # find what columns the names of the variables correspond to
  var_num <- match(v, names(d))
  var_y <- match("mass", names(d))
  # set starting learning rate (usually between 0.01 and 0.001)
  lr <- 0.001 # this was 0.01 in the  nasc version but here with biomass with 0.01 I tended to have target.trees < 1000 for several rounds of the loop

  target.trees <- 0 # resets the target trees , make sure set condition to lower than 1000

  while (target.trees < 1000) {
  
    mod.brt <- gbm.step(d,
                     gbm.x = var_num, # Your predictor variables, referring to columns in your data frame
                     gbm.y = var_y, # Your response variable
                     family = fam,
                     n.trees = 50,
                     bag.fraction = 0.8, # 0.75 is typical for larger dataset, but can increase bag fraction if dataset is small 
                     plot.main = F,   
                     learning.rate = lr,
                     tree.complexity = 2,
                     verbose = F)  
  
    if (object.size(mod.brt) > 0) {    # > 0 = good, check if hasn't crashed
      target.trees <- mod.brt$gbm.call$best.trees
    }
    else {
      target.trees <- 0  #check if model size is zero 
    }
    lr <- lr / 2
  }
  out <- tibble(model = list(mod.brt),
                lr = lr)
  return(out)
}

# create function to make dataframe predictions from a hierarchical modelling approach: binomial and gamma models
# to specific periods stored in env layers ras
# m_name is the model name = predictors chosen
# here fam should be a vector of the two families bin and logmass
# method is to indicate if the model is a GAM or BRT
Fun_Hpredras <- function(m_name, fam, m_tb, ras, met){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep <- tb %>% 
      filter(model_name == m_name & family == fam & method == met) %>% 
      pull(model_var)
    # model of presence
    m1 <- tb %>% 
      filter(model_name == m_name & family == fam[1] & method == met) %>% 
      pull(model)
    load(m1[[1]])
    if(met == "BRT"){m1m <- mod_brt$model[[1]]}
    if(met == "GAM"){m1m <- mod_gam}
    # model of abundance
    m2 <- tb %>% 
      filter(model_name == m_name & family == fam[2] & method == met) %>% 
      pull(model)
    load(m2[[1]])
    if(met == "BRT"){m2m <- mod_brt$model[[1]]}
    if(met == "GAM"){m2m <- mod_gam}
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week <- r_week[[var_tokeep[[1]]]]
            # then make predictions on raster!
            if(met == "BRT"){
              pred1 <- raster::predict(r_week, m1m, type = "response",
                                      n.trees = m1m$gbm.call$best.trees)
              pred2 <-  raster::predict(r_week, m2m, type = "response",
                                      n.trees = m2m$gbm.call$best.trees)
              pred2 <- exp(pred2)} # response is logged in brt
            if(met == "GAM"){
              pred1 <- raster::predict(r_week, m1m, type = "response")
              pred2 <- raster::predict(r_week, m2m, type = "response")
            }
            predH <- pred1 * pred2
            return(predH)
          }))
          # average across four weeks with NA.rm
          mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
          return(mean_pred)
      }))
  })
}

## BRT bin x GAM abun
Fun_crossENSpredras <- function(m_name, m_tb, ras){
  ddply(m_tb, ~folds, function(tb){
    # select env variables from model
    var_tokeep_bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "GAM") %>% 
      pull(model_var)
    var_tokeep_abu <- tb %>% 
      filter(model_name == m_name & family == "logmass" & method == "GAM") %>% 
      pull(model_var)
    # model of presence
    m.bin <- tb %>% 
      filter(model_name == m_name & family == "bin" & method == "BRT") %>% 
      pull(model)
    load(m.bin[[1]])
    m.bin.m <- mod_brt$model[[1]]
    # model of abundance
    m.abu <- tb %>% 
      filter(model_name == m_name & family == "logmass" & method == "GAM") %>% 
      pull(model)
    load(m.abu[[1]])
    m.abu.m <- mod_gam
    # produce predictions per week then average
    ras %>% 
      mutate(pred = map(env_ras, function(r){
        pred <- r %>% 
          mutate(pred_week = map(data_pred, function(r_week){
            # calculate log depth (suppres warnings because there are some positive values on the eastern edge of the extent that produce a warning with log(negative))
            suppressWarnings(r_week$logDEPTH <- calc(r_week$DEPTH, function(x) log(-x)))
            # get rid of variables in raster that are not in the model
            r_week_bin <- r_week[[var_tokeep_bin[[1]]]]
            r_week_abu <- r_week[[var_tokeep_abu[[1]]]]
            # then make predictions on raster!
            pred1 <- raster::predict(r_week_bin, m.bin.m, type = "response",
                                      n.trees = m.bin.m$gbm.call$best.trees)
            pred2 <- raster::predict(r_week_abu, m.abu.m, type = "response")
            predH <- pred1 * pred2
            return(predH)
        }))
        # average across four weeks with NA.rm
        mean_pred <- calc(stack(pred$pred_week), function(x) mean(x, na.rm = T))
        return(mean_pred)
      }))
  })
}

# create function to turn tibble of rasters into one dataframe that can generate a ggplot
Fun_meanpredmaps <- function(ras_tb){
  
  # transform into sf format
  d <- ras_tb %>% 
    mutate(pred_df = map(meanpred, function(r){
      # turn raster to sf object
      d <- r %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
      # add coordinates to dataframe of predictions
      d[c("x", "y")] <- st_coordinates(d)
      return(d)
    })) %>% 
    mutate(month = lubridate::month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE, locale="EN-us"),
           year = lubridate::year(as.POSIXct(pred_months, format = "X%Y.%m.%d")),
           period = paste(month, year)) %>%  
    dplyr::select(period, month, year, pred_df) %>%
    # arrange and mutate factor levels so that plots show up in the right order by months 
    arrange(year, month) %>% 
    mutate(period = factor(period, levels = unique(period))) %>% 
    unnest(pred_df)
  
  # create facetted maps by year in columns and months in rows
  g <- ggplot() +
    geom_tile(data = d, aes(x, y, fill = log(fit + 0.1))) +
    scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \nlog biomass",
                         n.breaks = 6) +  
    geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
    geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
    facet_wrap(~period, ncol = n_distinct(d$year), dir = "v") +
    coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
    mon_theme +
    xlab("") +
    ylab("") +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
    
    return(g)
}

# a derivative of the function meanpredmaps but applied to biomass proportions by species
Fun_propmaps <- function(ras_tb){
  
  # transform into sf format
  d <- ras_tb %>% 
    mutate(pred_df = map(meanpred, function(r){
      # turn raster to sf object
      d <- r %>% 
        st_as_stars() %>%
        st_as_sf(as_points = T, crs = 32610) %>% 
        rename(fit = layer)
      # add coordinates to dataframe of predictions
      d[c("x", "y")] <- st_coordinates(d)
      return(d)
    })) %>% 
    mutate(month = lubridate::month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE, locale="EN-us"),
           year = lubridate::year(as.POSIXct(pred_months, format = "X%Y.%m.%d")),
           period = paste(month, year)) %>%  
    dplyr::select(period, month, year, pred_df) %>%
    # arrange and mutate factor levels so that plots show up in the right order by months 
    arrange(year, month) %>% 
    mutate(period = factor(period, levels = unique(period))) %>% 
    unnest(pred_df)
  
  # create facetted maps by year in columns and months in rows
  g <- ggplot() +
    geom_tile(data = d, aes(x, y, fill = fit*100)) +
    scale_fill_viridis_b(na.value = "transparent", 
                         name = "Predicted \n Proportion (%)",
                         n.breaks = 6) +  
    geom_sf(data = iso_sf_utm[iso_sf_utm$g_2020_ %in% c(-200, -1500), ], 
            col = "grey80", linewidth = 0.1) +
    geom_sf(data = coast_sf_utm, col = NA, fill = "grey10") +
    facet_wrap(~period, ncol = n_distinct(d$year), dir = "v") +
    coord_sf(xlim = c(-30592.78, 561873.5), ylim = c(4150571, 5378175), expand = F) +
    mon_theme +
    xlab("") +
    ylab("") +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5),
          axis.text.x = element_text(hjust = 0.9),
          text=element_text(face="bold", size=8),
          legend.key.size = unit(0.8, "lines"),
          panel.spacing = unit(0.5, "lines"),
          plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
    
    return(g)
}
```

# Running SDMs
## Set up models

### Predictor formulas
```{r}
# only the main candidate models are being run in this code to speed up the process. See "Github_all_models" for codes of all models presented in the manuscript
form_chr <- list(
     "topographic_logdep" = " mass ~
                   s(logDEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5)",
     "dynamic_logdep" = "mass ~ 
                   s(logDEPTH, k = 5) +
                   s(DIS_CANYON, k = 3) + 
                   s(SLOPE, k = 3) + 
                   s(sst, k = 3) + 
                   s(ssh, k = 3) + 
                   s(sshsd, k = 3) + 
                   s(sstsd, k = 3) + 
                   s(eke, k = 3) +
                   s(curl, k = 3) + 
                   s(ild_05, k = 3) + 
                   s(bbv_200, k = 3)",
     "dynamic_sst" = "mass ~
                   s(logDEPTH, k = 5) + 
                   s(SLOPE, k = 5) + 
                   s(DIS_CANYON, k = 5) + 
                   s(sst, k = 5)")

model_type <- c(rep("topo", 1), rep("dyn", 2))
model_var <- list(
                   c("logDEPTH","SLOPE", "DIS_CANYON"),
                   c("logDEPTH","DIS_CANYON", "SLOPE", "sst", "ssh", "sshsd", "sstsd","eke", "curl", "ild_05", "bbv_200"),
                   c("logDEPTH","SLOPE", "DIS_CANYON", "sst"))
```

### Folds and model types
```{r}
# because there are at least two cruises in each season (Feb, May, Sep) we can leave one out at each run and still have the three seasons represented in the training dataset
Fun_folding <- function(df){
  # set seed to NULL so that we can rerun multiple times until we get a combination of folds without zero presence
  folds <- splitTools::create_folds(df$CruiseID, k = 10, seed = 1, type = "grouped")
  d_cv <- llply(folds, function(f){
    df$fold <- "test"
    df[f,]$fold <- "train"
    return(df)
  })
  return(d_cv)
}

unique(biomass_df$CruiseID)

# there are ten folds in the bongo biomass dataset (vs 8 in the nasc) because the 2018 february/march cruise does not have nasc data
krillmass_cv <- tibble(folds = paste0("Fold", c(1:10)),
                   data = Fun_folding(biomass_df))

modelsH_krillmass_tb <- tibble(model_form = form_chr) %>%
    mutate(model_name = names(form_chr),
         model_type = model_type,
         model_var = model_var,
         model_form = map(model_form, as.formula),
         data = rep(list(krillmass_cv), length(form_chr))) %>% 
  unnest(cols = c(data))
modelsH_krillmass_tb

# concatenate the tibble 2 times for the two krill species
modelsH_krillmass_tb <- rbind(modelsH_krillmass_tb, modelsH_krillmass_tb) %>% 
  mutate(group = rep(c("EPAC", "TSPIN"), each = nrow(modelsH_krillmass_tb))) %>% 
  # create a mass variable by species
  mutate(data = pmap(.l = list(x = data, y = group), .f = function(x, y){
    # add the new variable by sp
    x$mass <- as.data.frame(x[,paste0("mass_", y)])[, 1]
    return(x)
  }))
```

### Run models
```{r, eval = F}
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  ## BINOMIAL MODELS
  mutate(bin = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, gr = group, fo = folds), .f = function(a, b, c, mn, gr, fo){ 
      # create data set with binomial response
      a_modified <- a %>% mutate(mass = ifelse(mass == 0, 0, 1))
      # in the case of GAMs, the explanatory variables should not have NAs
      a_modified_gam <- a_modified[complete.cases(a_modified[b]), ]
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified, fold == "train"), 
                           v = b, fam = "bernoulli")
      mod_gam <- gam(c, family = binomial(link = "logit"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_bin_", gr, "_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_bin_", gr, "_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return models in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    })) %>%
  ## POSITIVE VALUES MODEL
  mutate(logmass = pmap(list(a = data, b = model_var, c = model_form, mn = model_name, gr = group, fo = folds), .f = function(a, b, c, mn, gr, fo){ 
      # create data set with positive biomass response
      a_modified_gam <- a %>% 
                    filter(mass > 0)
      a_modified_gam <- a_modified_gam[complete.cases(a_modified_gam[b]), ]
      # in the case of BRT the log should be applied prior to running the model
      a_modified_brt <- a %>% 
                    filter(mass > 0) %>% 
                    mutate(mass = log(mass)) # apply a log link
      
      print(paste0("m_abu_", gr, "_", mn, "_", fo))    
      # run models
      mod_brt <- Fun_mybrt(d = subset(a_modified_brt, fold == "train"), 
                           v = b, fam = "gaussian")
      mod_gam <- gam(c, family = Gamma(link = "log"), 
                      method = "REML",
                      select = T,
                      data = subset(a_modified_gam, fold == "train"))
      
      # save models on disk
      path_brt <- paste0("./Outputs/models/m_brt_logmass_", gr, "_", mn, "_", fo, ".RData")
      save(mod_brt, file = path_brt)
      path_gam <- paste0("./Outputs/models/m_gam_logmass_", gr, "_", mn, "_", fo, ".RData")
      save(mod_gam, file = path_gam)
      
      rm(mod_gam); rm(mod_brt)
      
      # return path to model objects in a tibble
      return(tibble(method = c("BRT", "GAM"),
                    model = list(path_brt, path_gam)))
    
    })) 

# pivot tibble and calculate deviance explained by each model
# in training and in test
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  pivot_longer(cols = c(bin, logmass), names_to = "family", values_to = "mod") %>% 
  unnest(cols = c(mod))

# correct data depending on family and method
modelsH_krillmass_tb <- modelsH_krillmass_tb %>%
  mutate(data = pmap(list(d = data, v = model_var, fa = family, m = method), 
                     function(d, v, fa, m){
      if(fa == "bin" & m == "BRT"){
        dd <- d %>% mutate(mass = ifelse(mass == 0, 0, 1))
      }
      if(fa == "logmass" & m == "BRT"){
        dd <- d %>% filter(mass > 0) %>% mutate(mass = log(mass))
      }
      if(fa == "bin" & m == "GAM"){
        dd <- d %>% mutate(mass = ifelse(mass == 0, 0, 1))
        if(!is.null(v)){dd <- dd[complete.cases(dd[v]), ]}
      }
      if(fa == "logmass" & m == "GAM"){
        dd <- d %>% filter(mass > 0)
      }
      return(dd)
  }))

save(modelsH_krillmass_tb, file = "./Outputs/modelsH_krillmass_tb.RData")
```

# Model performance

## Performance by method by family
```{r, warning = F, message = F, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")

modelsH_krillmass_perf <- modelsH_krillmass_tb %>%
  mutate(dev.exp = pmap_dbl(list(mo = model, me = method), function(mo, me){
    load(mo)
    if(me == "BRT"){
      dev <- round((mod_brt$model[[1]]$self.statistics$mean.null - mod_brt$model[[1]]$cv.statistics$deviance.mean)/mod_brt$model[[1]]$self.statistics$mean.null*100, 1)
    }
    if(me == "GAM"){
      dev <- round(summary(mod_gam)$dev.expl*100, 1)
    }
    rm(mo)
    return(dev)
  })) %>%
  mutate(AUC.train = pmap_dbl(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    pred.train <- predict(m, subset(d, fold == "train"), type = "response")
    if(fa == "bin"){
        auc <- round(pROC::roc(subset(d, fold == "train")$mass, pred.train)$auc, 3)
    } else {auc <- NA}
    return(auc)
  }))


modelsH_krillmass_perf <- modelsH_krillmass_perf %>%
  # For the test fold, we calculate either AUC for the bin model, or pear.cor/spea.cor for the logmass model
  mutate(metrics.test = pmap(list(mo = model, fa = family, d = data, me = method), function(mo, fa, d, me){
    load(mo)
    if(me == "BRT"){m <- mod_brt$model[[1]]}
    if(me == "GAM"){m <- mod_gam}
    test.data <- subset(d, fold == "test")
    test.data$pred <- predict(m, test.data, type = "response")
    if(fa == "bin"){ # binomial models need zeros and ones
      if(nrow(subset(test.data, !is.na(pred) & mass > 0)) > 1 &
        nrow(subset(test.data, !is.na(pred) & mass == 0)) > 1){
          auc <- round(pROC::roc(test.data$mass, test.data$pred)$auc, 3)
          spea.cor <- NA
          pear.cor <- NA
      } else { spea.cor <- NA ; auc <- NA ; pear.cor <- NA}
    } else { # abundance models (no zeros)
      if(nrow(subset(test.data, !is.na(pred))) > 2){
          auc <- NA
          spea.cor <- round(cor.test(test.data$mass, test.data$pred, method = "spearman")$estimate, 3)
          pear.cor <- round(cor.test(test.data$mass, test.data$pred, method = "pearson")$estimate, 3)
      } else { spea.cor <- NA ; auc <- NA ; pear.cor <- NA}
    }
    metrics <- tibble(spea.cor = spea.cor, pear.cor = pear.cor, AUC.test = auc)
    return(metrics)
  })) %>% 
  unnest(cols = c(metrics.test)) %>% 
  dplyr::select(group, model_name, family, method, folds, dev.exp, AUC.train, AUC.test, pear.cor, spea.cor) %>% 
  arrange(group, family)

save(modelsH_krillmass_perf, file = "./Outputs/modelsH_krillmass_perf.RData")
```


```{r, warning = F, message = F}
load("./Outputs/modelsH_krillmass_perf.RData")

(mean_perf <- modelsH_krillmass_perf %>% 
  group_by(group, model_name, family, method) %>% 
  summarize(mean.dev.exp = mean(dev.exp, na.rm = T),
            mean.AUC.train =  mean(AUC.train, na.rm = T),
            mean.AUC.test = mean(AUC.test, na.rm = T),
            mean.pear.cor = mean(pear.cor, na.rm = T),
            mean.spea.cor = mean(spea.cor, na.rm = T)) %>% 
  mutate_if(is.numeric, ~ round(., digits = 3)) %>% 
  select(group, family, everything()) %>% 
  arrange(group, family, -mean.dev.exp))

```

# Predicted maps of biomass

Selected topographic logdepth of model

```{r, eval = F}
load("./Outputs/modelsH_krillmass_tb.RData")
     
modelsH_krillmass_topolog <- modelsH_krillmass_tb %>% 
  filter(model_name == "topographic_logdep")

save(modelsH_krillmass_topolog, file = "./Outputs/modelsH_krillmass_topolog.RData")
```

Months x years to predict year-round
```{r}
# select combination of months and years of interest
pred_subset <- tidyr::crossing(years = c(2018:2022), months = c(1:12))
pred_subset <- paste0("X", pred_subset$years, ".", pred_subset$months)

# select only these periods in the full environmental data rasters                   
env_pred_yearround <- env_pred_full %>% 
  filter(substr(pred_months, 1, 7) %in% pred_subset)

# check that we have the right months represented
unique(env_pred_yearround$pred_months)
```

Generate predictions from the Ensemble model
```{r, eval = F}
load("./Outputs/modelsH_krillmass_topolog.RData")

predyear_crossENS_topolog_mass <- ddply(modelsH_krillmass_topolog, .(group), function(d) {
  Fun_crossENSpredras(m_name = "topographic_logdep",  m_tb = d, ras = env_pred_yearround)})
save(predyear_crossENS_topolog_mass, file = "./Outputs/predyear_crossENS_topolog_mass.RData")

predyear_crossENSavg_topolog_mass <- ddply(predyear_crossENS_topolog_mass, .(group, pred_months), function(d){
  tibble(meanpred = list(calc(stack(d$pred), function(x) mean(x, na.rm = T))))
})
save(predyear_crossENSavg_topolog_mass, file = "./Outputs/predyear_crossENSavg_topolog_mass.RData")
```


# Relationships

With selected topographic-logdepth model

## Variable influence
```{r}
load("./Outputs/modelsH_krillmass_topolog.RData")

## GAM
Fun_VarPVGAM <- function(m){
  load(m)
  data.frame(predvar = names(summary(mod_gam)$chi.sq),
             pv = summary(mod_gam)$s.pv)
}

Fun_VarInfluGAM <- function(d){
  data.frame(count0.001 = length(d$pv[d$pv <= 0.001]),
             count0.01 = length(d$pv[d$pv <= 0.01]),
             count0.05 = length(d$pv[d$pv <= 0.05]),
             mean_pvalue = round(mean(d$pv), 3))
}

influ_gam_topolog_mass <- modelsH_krillmass_topolog %>% 
  filter(method == "GAM") %>% 
  mutate(influence = map(model, function(m) Fun_VarPVGAM(m))) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  ddply(., .(group, method, family, predvar), function(d) Fun_VarInfluGAM(d))
influ_gam_topolog_mass
  
save(influ_gam_topolog_mass, file = "./Outputs/influ_gam_topolog_mass.RData")

### BRT
influ_brt_topolog_mass <- modelsH_krillmass_topolog %>% 
  filter(method == "BRT") %>% 
  mutate(influence = map(model, function(m) {
    load(m)
    mod_brt$model[[1]]$contributions})) %>% 
  select(group, method, family, folds, influence) %>% 
  unnest(cols = c(influence)) %>% 
  group_by(group, method, family, var) %>% 
  summarize(meanInf = mean(rel.inf),
            sdInf = sd(rel.inf)) %>% 
  rename(predvar = var)
influ_brt_topolog_mass
save(influ_brt_topolog_mass, file = "./Outputs/influ_brt_topolog_mass.RData")

```

## Partial Dependence plots

With selected topographic-logdepth model and ensemble approach

```{r}
# create a dataframe filled with the average of each variable
suppressWarnings(
mean_pred_df <- biomass_df %>% 
    dplyr::select(modelsH_krillmass_topolog$model_var[[1]]) %>% 
    summarise_if(is.numeric, function(x) rep(mean(x, na.rm = T), 200))
)

# for depth I am fixing the mean to 200 m instead of the mean which is about 400 m
mean_pred_df$logDEPTH <- log(200)

# run predictions over model type and folds
ENScurves_krillmass_topolog <- modelsH_krillmass_topolog %>% 
  filter((method == "BRT" & family == "bin") | (method == "GAM" & family == "logmass")) %>% 
  mutate(pred = map2(model, method, function(m, me){
    load(m)
    if(me == "BRT"){
      m <- mod_brt$model[[1]]
      var_selec <- m$contributions %>% 
        filter(rel.inf > 5)
    }
    if(me == "GAM"){
      m <- mod_gam
      var_selec <- data.frame(predvar = names(summary(m)$chi.sq),
                        pv = summary(m)$s.pv) %>% 
        filter(pv <= 0.05) %>% 
        mutate(var = substr(predvar, 3, (nchar(predvar)-1)))
    }
    new_pred <- ddply(var_selec, ~var, function (v){ # for each predictor, calculate response with model m
        extremes <- quantile(biomass_df[v$var[1]][,1], probs = c(0.01, 0.99), na.rm = T)
        mean_pred_x_df <- mean_pred_df
        mean_pred_x_df[,v$var[1]] <- seq(extremes[1], extremes[2], length.out = 200)
        if(me == "BRT"){
          pred <- data.frame(fit = raster::predict(m, mean_pred_x_df, type = "response",
                                      n.trees = m$gbm.call$best.trees))
          pred$se.fit <- NA # no error estimate in the BRT
        }
        if(me == "GAM"){
          pred <- dismo::predict(m, mean_pred_x_df, type = "response", se.fit = T)
        }
        new_pred_df <- data.frame(x = mean_pred_x_df[, v$var[1]])
        names(new_pred_df) <- "x"
        new_pred_df$predvar <- v$var[1]
        new_pred_df$fit <- pred$fit
        new_pred_df$se <- pred$se.fit
      return(new_pred_df)
    })
  }))

ENScurves_krillmass_topolog <- ENScurves_krillmass_topolog %>% 
  select(group, family, folds, pred) %>% 
  unnest(cols = c(pred))

save(ENScurves_krillmass_topolog, file = "./Outputs/ENScurves_krillmass_topolog.RData")
```


# Predicted species-scaled NASC

## Predicted biomass proportion maps

Calculate the proportion of EPAC and TSPIN over monthly rasters
```{r}
load("./Outputs/predyear_crossENSavg_topolog_mass.RData")

pred_propbiomass_ENStopolog <- ddply(predyear_crossENSavg_topolog_mass, .(pred_months), function(d){
  tibble(epac_prop = list(d$meanpred[[1]]/(d$meanpred[[1]] + d$meanpred[[2]])),
  tspin_prop = list(d$meanpred[[2]]/(d$meanpred[[1]] + d$meanpred[[2]])))
})
```

Combine predictions of proportions into a single object. This code was initially built to integrate several different model outputs and compare the fit between predicted and observed species proportions. In this version of code, we simplified the loop to only include our selected model: topographic_logdepth.
```{r}
pred_propbiomass <- pred_propbiomass_ENStopolog
names(pred_propbiomass) <- c("pred_months", "epac_ENStopolog", "tspin_ENStopolog")

pred_propbiomass <- pred_propbiomass %>% 
  mutate(ras = pmap(list(eetl = epac_ENStopolog, tetl = tspin_ENStopolog), 
                    function(eetl, tetl){
                           r <- stack(list(eetl, tetl))
                           names(r) <- c("epac_ENStopolog", "tspin_ENStopolog")
                           return(r)
                         })) %>% 
  mutate(year.month = paste(str_split_i(pred_months, "[.]", 1), 
                            str_split_i(pred_months, "[.]", 2), sep = ".")) %>% 
  select(year.month, ras)

save(pred_propbiomass, file = "./Outputs/pred_propbiomass.RData")
```

## Calculate NASC x Biomass proportions per species

```{r}
################################################################
########## ENS NASC dynlogdep x ENS Biomass prop Topolog #######
pred_nascXmass_ENSENStopolog <- join(predyear_crossENSavg_dynlogdep, pred_propbiomass_ENStopolog, by = "pred_months") %>% 
  mutate(pred_epac = pmap(list(nasc_pred = meanpred, prop_pred = epac_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) %>% 
  mutate(pred_tspin = pmap(list(nasc_pred = meanpred, prop_pred = tspin_prop), function(nasc_pred, prop_pred){
    nasc_pred * prop_pred
  })) 

save(pred_nascXmass_ENSENStopolog, file = "./Outputs/pred_nascXmass_ENSENStopolog.RData")
```

Map predictions for three months of interest
```{r}
pred_nascXmass_ENSENStopolog_selec <- pred_nascXmass_ENSENStopolog %>% 
  filter(substr(pred_months, 7, 7) %in% c("2", "5", "9")) %>% 
  filter(as.numeric(substr(pred_months, 2, 5)) >= 2018 &
           as.numeric(substr(pred_months, 2, 5)) <= 2022)

g_epac <- pred_nascXmass_ENSENStopolog_selec %>% 
  mutate(meanpred = pred_epac) %>% 
  Fun_meanpredmaps()

g_epac + ggtitle("EPAC: pred NASC dynlogdep ENS x pred biomass proportion topolog ENS")

g_tspin <- pred_nascXmass_ENSENStopolog_selec %>% 
  mutate(meanpred = pred_tspin) %>% 
  Fun_meanpredmaps()

g_tspin + ggtitle("TSPIN: pred NASC dynlogdep ENS x pred biomass proportion topolog ENS")
```

## Evaluate proportion predictions

Create function to generate plots of observed vs predicted proportion
```{r}
Fun_evalprop <- function(d, x, y, method, sp){
  d$x2 <- unlist(d[x]/d$mass_tot)
  ggpubr::ggscatter(d, x = "x2", y= y, add = "reg.line") +
    ggpubr::stat_cor(label.y = 1) +
    ggpubr::stat_regline_equation(label.y = 0.8) +
    mon_theme +
    ggtitle(method) +
    xlab(paste0("Observed proportion of ", sp)) +
    ylab(paste0("Predicted proportion of ", sp)) +
    ylim(0, 1) +
    xlim(0, 1)
}
```

Create dataset of evaluation with NH line and NCC stations
```{r}
biomass_nhl_df$program <- "NHL"
biomass_df$program <- "NCC"

all_biomass_eval_df <- rbind(biomass_nhl_df %>% 
             select(program, date_raster_week, Station, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, utmx, utmy),
                             biomass_df %>% 
             select(program, date_raster_week, Station, Sample.Date, mass_EPAC, mass_TSPIN, mass_tot, utmx, utmy)) %>% 
  mutate(year.month = paste(str_split_i(date_raster_week, "[.]", 1), str_split_i(date_raster_week, "[.]", 2), sep = ".")) %>% 
  ddply(., ~year.month, function(d){ # divide biomass data by month
    # select predictions for that month
    pred_propbiomass_selec <- pred_propbiomass %>% filter(year.month == d$year.month[1])
    # extract epac and tspin proportions calculated with Ensemble topographic logdepth models
    d[c("epac_ENStopolog", "tspin_ENStopolog")] <- raster::extract(x = pred_propbiomass_selec$ras[[1]], y = d[c("utmx", "utmy")])
    return(d)
  }) %>% 
  filter(mass_tot != 0) # remove stations at which there was absolutely no krill (15 stations in NH line)

save(all_biomass_eval_df, file = "./Outputs/all_biomass_eval_df.RData")
```

Plot regression between predicted and observed biomass proportions
```{r}
a <- Fun_evalprop(all_biomass_eval_df, x = "mass_TSPIN", y = "tspin_ENStopolog", method = "ENS - topolog", sp = "Tspin")
b <- Fun_evalprop(all_biomass_eval_df, x = "mass_EPAC", y = "epac_ENStopolog", method = "", sp = "Epac")

a + b
```

# Trends in species-scaled NASC predictions

## Temporal trends across the whole study region

```{r}
load("./Outputs/pred_nascXmass_ENSENStopolog.RData")

krill_final_ENSxENStopolog_trend <- pred_nascXmass_ENSENStopolog %>%
  mutate(sum_abu_epac = map_dbl(pred_epac, function(r){
    sum(values(r), na.rm = T)
  })) %>%
  mutate(sum_abu_tspin = map_dbl(pred_tspin, function(r){
    sum(values(r), na.rm = T)
  })) %>% 
  dplyr::select(pred_months, sum_abu_epac, sum_abu_tspin) %>% 
  pivot_longer(cols = c(sum_abu_epac, sum_abu_tspin), 
               names_to = "group", values_to = "sum_abu",
               names_prefix = "sum_abu_") %>% 
  mutate(month = lubridate::month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE, locale="EN-us"),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_final_ENSxENStopolog_trend, file = "./Outputs/krill_final_ENSxENStopolog_trend.RData")
```

Mean proportion of Tspin and Epac
```{r}
prop_sp <- krill_final_ENSxENStopolog_trend %>% 
  group_by(group) %>% 
  summarize(mean_abu = mean(sum_abu))

prop_sp[prop_sp$group == "tspin", "mean_abu"] / prop_sp[prop_sp$group == "epac", "mean_abu"]
```

## Temporal trends along NH Line

Create a 5 km high raster that just covers the NH line
```{r}
# mean latitude of NH Line
nhl_lat <- mean(biomass_nhl_df$utmy)
# creating a 5 km band around that latitude
extent_nhl_band <-  extent(extent(depth_ras)@xmin, extent(depth_ras)@xmax, nhl_lat - 2500, nhl_lat + 2500)

# extract depth along the same line
depth_nhl <- crop(depth_ras, extent_nhl_band)
```

Extract predicted species-scaled NASC in the NH Line raster band
```{r}
load("./Outputs/pred_nascXmass_ENSENStopolog.RData")

# extract as dataframe the abundance pred for epac and tspin along that band
krill_nhl_pred_ENSENStopolog_trend <- pred_nascXmass_ENSENStopolog %>%
  mutate(sum_nhl_epac = map(pred_epac, function(r){
    names(r) <- "pred_epac"
    pred_nhl <- crop(r, extent_nhl_band)
    s <- stack(list(depth_nhl, pred_nhl))
    d <- as.data.frame(s)
    d[c("utmx", "utmy")] <- coordinates(s)
    return(d)
  })) %>%
  mutate(sum_nhl_tspin = map(pred_tspin, function(r){
    names(r) <- "pred_tspin"
    pred_nhl <- crop(r, extent_nhl_band)
    d <- as.data.frame(pred_nhl)
    return(d)
  })) %>% 
  dplyr::select(pred_months, sum_nhl_epac, sum_nhl_tspin) %>% 
  unnest(cols = c(sum_nhl_epac, sum_nhl_tspin)) %>% 
  filter(!is.na(DEPTH)) %>% # remove portion of the band that is on land
  pivot_longer(cols = c(pred_epac, pred_tspin), 
               names_to = "group", values_to = "pred",
               names_prefix = "pred_") %>% 
  mutate(month = lubridate::month(as.POSIXct(pred_months, format = "X%Y.%m.%d"),
                        label = TRUE, abbr = TRUE, locale="EN-us"),
           year = as.factor(year(as.POSIXct(pred_months, format = "X%Y.%m.%d"))),
           period = paste(month, year)) %>% 
  arrange(group, year, month) %>% 
  mutate(period = factor(period, levels = unique(period)))

save(krill_nhl_pred_ENSENStopolog_trend, file = "./Outputs/krill_nhl_pred_ENSENStopolog_trend.RData")
```

